### 2022-01-11 proposal/research_proposal.Rmd
after work on delft device

---
title: "Overlooked dynamics of algorithmic recourse"
subtitle: "Research Proposal"
author:
- "Patrick Altmeyer (p.altmeyer@tudelft.nl)"
date: "`r format(Sys.Date(), '%B, %Y')`"
output: 
  bookdown::pdf_document2: 
    toc: false
    includes:
      in_header: ../tex/preamble.tex
  bookdown::html_document2: default
fontsize: 12pt
bibliography: ../bib.bib
link-citations: true
abstract: "..."
---

```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(
  echo = FALSE,
  warning=FALSE,
  message=FALSE,
  out.width = 450,
  fig.align = "center"
)
```

# Introduction

In their NeurIPS 2021 paper @upadhyay2021towards demonstrate that existing approaches to algorithmic recourse are not robust to exogenous domain and model shifts. A related and so far unexplored idea is that recourse itself may lead to endogenous domain and model shifts, if a substantial share of individuals acts on it. We believe that this can have implications on the validity and cost of algorithmic recourse that have so far been overlooked.

# Related work

## @wachter2017counterfactual

## @schut2021generating

## @upadhyay2021towards

## Other

- @joshi2019towards
- @ustun2019actionable
- @karimi2021algorithmic

# Preliminary findings

## Synthetic data

## German credit data