### 2022-02-03 students/project_counterfactual_generators.qmd
synthetic and boston housing experiments done. Finished student project proposal.

---
title: "Endogenous model shifts in algorithmic recourse"
author: "Patrick Altmeyer"
date: "February 03, 2022"
format: gfm
bibliography: ../bib.bib
---

Responsible Professor: Dr. Cynthia Liem
Supervisor: Patrick Altmeyer

## Project overview

Algorithmic recourse is an approach towards not only explaining black-box machine learning algorithms, but provide subjects 
who have been adversely affected by such with a realistic set of actions they can take in order to revise their outcome. 
The approach has been quickly growing in popularity during recent years, perhaps because the underlying tend to be methodologies tend to be intuitive and easy to implement. This project aims to shed some light on what happens *after* recourse is actually implemented. In particular, we want to investigate if and how the resulting domain shifts lead to model shifts and compare outcomes for different recourse generators.

## Motivation and background

The topic is also of great importance to society since automated decision-making through black-box algorithm affects all of us - inlcuding you! Suppose, for example, that upon graduation you apply for a job at a large coorperation. Unfortunately, you soon receive an email stating that you have not "met the short-listing criteria and no further feedback can be provided". Unbeknown to you, the company has used an algorithm to decide who makes it onto the short list. The HR teams does not fully understand the inner workings of the algorithm, so they genuinely cannot provide you with feedback.

Algorithmic recourse aims to resolve these types of issues. But existing work has paid little attention to the dynamics of 
recourse. The first paper to consider the problem that algorithms are updated in response to (exogenous) domain shifts was presented at NeurIPS in December 2021 [@upadhyay2021towards]. The authors propose a recourse generator that is robust to "small (and not drastic)" shifts. But rather than small shifts, we find that the actual implementation of recourse can lead to potentially quite large domain shifts and consequently model shifts. The animation below illustrates this: after a small share of individuals moves across the decision boundary, the classifier is updated in response. Evidently, the model shifts are quite significant in this toy example. Below we refer to these types of shifts as *endogenous* shifts.

We think that endogenous shifts deserve attention for various reasons:

1. Our initial experimental results suggests that even generators that aim to produce realisitic counterfactuals can induce significant domain and model shifts.
2. Model shifts in response to some individuals' recourse may lead to undesired consequences (label switch) for other individuals.
3. Individuals that have implemented recourse may end up forming their own distinct cluster in the feature domain. The original classifier could therefore be adapted to distinguish "recourse" indviduals in the target class from those that were always in the target class.

![](../work/www/synthetic/bayesian.gif)

## Research questions

Collectively, students will discuss and agree on an answer to the following question:

- How can we quantify endogenous domain and model shifts?

Individually, each of the students will then choose one of the recourse generators already implemented in [CARLA](https://github.com/carla-recourse/CARLA) and run a simple experiment (details below) to generate endogenous shifts and quantify their magnitude.

Collectively, the students will finally tackle the following questions:

- Does the magnitude of induced model shifts vary by recourse generator? 
- If so, what factors might be playing a role here?
- Based on your findings, what appear to be good ways to deal mitigate endogenous shifts?

While the individual part should really be done individually, do keep in mind that you will end up comparing the results of your experiments and you should therefore all generate the same, comparable output.

## Methodology

Overall, this will be a project focused on benchmarking and experimental design: you will learn about how to use existing tools for benchmarking to asses outcomes from your own experiments.

For a gentle introduction to the topic of algorithmic recourse you can check out this [primer](https://towardsdatascience.com/individual-recourse-for-black-box-models-5e9ed1e4b4cc) written by Patrick as part of his PhD application. To familiarize yourself further with the topic and [CARLA](https://github.com/carla-recourse/CARLA) you should all read this [paper](https://arxiv.org/pdf/2108.00783.pdf).

You are free to choose any of the generators already implemented [CARLA](https://github.com/carla-recourse/CARLA), but below are five suggestions:

1. [Wachter](https://arxiv.org/ftp/arxiv/papers/1711/1711.00399.pdf)
2. [REVISE](https://arxiv.org/pdf/1907.09615.pdf)
3. [CEM](https://arxiv.org/pdf/1802.07623.pdf)
4. [CEM-VAE](https://arxiv.org/pdf/1802.07623.pdf)
5. [AR-LIME](https://arxiv.org/pdf/1809.06514.pdf)

Since the generators are all conveniently implemented in the same library, running the experiment should require roughly the 
same effort for all of them. While you are not expected to fully understand in detail the methodology underlying your
chosen generator, you should study the relevant paper and ask yourself: what makes your generator particular? What assumptions do the authors make? How do you expect these particularities to affect recourse outcome with respect to endogenous shifts?

**Experiment**

The idea is that you replicate a variation of the following experiement already implemented by Patrick:

1. Train an algorithm $\mathcal{M}$ for a binary classification task. 
2. Determine a target class. Generate recourse for a share $\mu$ of randomly selected individuals in the other class to revise their label (i.e. move to the target class). Store some of the conventional benchmark measures readily implemented in CARLA (cost, success rate, ...)
3. Implement recourse for those indviduals and quantify the domain shift.
4. Retrain $\mathcal{M}$ and quantify the model shift.
5. Repeat steps 1-4 for $K$ rounds.

You are free to either follow Patrick's recipe exactly or tweak the various parameters that determine the experiemntal design as you see fit. Make sure that you end up running the exact same experiment though, in order for the results to be comparable.

**Hypothesis**

- I expect to find that generators that focus on creating realisitic counterfactuals in high-density regions of the feature domain (REVISE, CEM-VAE) tend to be associated with endogenous shifts of a smaller magnitude compared to generators that do not specifically address this (CEM, Wachter, AR-LIME).

## Relation between sub-projects

The students will collectively design and establish an evaluation benchmark, while each focusing on a particular framework implementation to ultimately evaluate in the benchmark. Researching the intricacies of each framework will lead to clearly individual contributions, while part of the relevant literature is shared, and the final comparison connects the individual insights.

## Responsible Engineering / Computer Science / Research

As noted above, algorithmic recourse was designed in the context of responsible machine learning. Work on this topic should foster an awareness that automated decision-making can have detrimental consequences for individuals and that many researchers are working on approaches to deal with some of those consequences. Comparing different approaches to the issue will also make it clear that in the context of research there is rarely (if ever) one right solution. More to the point of this particular research project, solutions to one particular issue usually open up new questions.

## Possible publication

In 2021 there was an [ICML workshop](https://icml.cc/Conferences/2021/ScheduleMultitrack?event=8363) on algorithmic recourse. Should this workshop take place again, the results of this project could lead to a poster presentation. Later in the year NeurIPS is coming up, which in 2021 also attracted conference and workshop papers on algorithmic recourse and benchmarking.

## References