### 2022-04-13 dev/work/german_credit.qmd

---
jupyter: julia-1.6
---

## German credit data

```{julia}
include("../src/load.jl")
using AlgorithmicRecourse
using CSV;
using DataFrames;
using LinearAlgebra;
```

```{julia}
df = CSV.read("../data/credit.csv", DataFrame);
df[!,:] = convert.(Float64, df[!,:]); # convert to floats
y = df[:,:y]
N = length(y)
X = Matrix(df[:,Not(:y)])';
X_train = copy(X)
y_train = copy(y);
```

```{julia}
using MLDataUtils, Random, Flux
Random.seed!(1234)

# # Preproc:
# X_train, y_train = oversample((X_train,y_train))
# # Standardize continuous features:
# using StatsBase
# dt = fit(ZScoreTransform, X_train[1:3,:], dims=1)
# X_train[1:3,:] = StatsBase.transform(dt, X_train[1:3,:])

# Stack:
xs = Flux.unstack(X_train,2)
ys = Flux.flatten(y_train)
data = zip(xs, ys)
nn = Models.build_model(input_dim=size(X_train)[1], n_hidden=32)
loss(x, y) = Flux.Losses.logitbinarycrossentropy(nn(x), y)
ps = params(nn);
```

```{julia}
using Flux.Optimise: update!, ADAM
using Statistics
opt = ADAM()
epochs = 50
avg_loss(data) = mean(map(d -> loss(d[1],d[2]), data))

using Plots
anim = Animation()
plt = plot(ylim=(0,avg_loss(data)), xlim=(0,epochs), legend=false, xlab="Epoch")
avg_l = []

for epoch = 1:epochs
  for d in data
    gs = gradient(params(nn)) do
      l = loss(d...)
    end
    update!(opt, params(nn), gs)
  end
  avg_l = vcat(avg_l,avg_loss(data))
  plot!(plt, avg_l, color=1, title="Average (training) loss")
  frame(anim, plt)
end

gif(anim)
```

**EVEN THE NEURAL NET JUST CLASSIFIES ALL AS 1, WHAT IS GOING ON HERE?**

```{julia}
sum(round.(Flux.σ.(nn(X_train)))' .== 1)
```

```{julia}
using EvalMetrics
using Plots
y_hat = predict(model,X)
rocplot(y,y_hat)
```

```{julia}
generators = (wachter = generate_recourse_wachter, schut = generate_recourse_schut);
generator_args = (wachter=(λ=0.01,), schut=(T=1000,δ=0.05));
target = 1;
```

```{julia}
props = [0.01,0.05,0.1,0.25]
results = DataFrame()
for prop in props
    results_prop = run_experiment(X,y,bayes_logreg,target,generators,generator_args,experiment_dynamic)
    insertcols!(results_prop, :prop => prop)
    results = vcat(results, results_prop)
end
```

Possible that these results are driven by the following fact:

- classifier is biased towards target class (try random over/undersampling)

```{julia}
dt_plot = groupby(results, [:period, :generator, :prop]) |>
    gdf -> combine(gdf, :validity .=> [mean, std] .=> [:mean, :std])
dt_plot[!,:ymin] = dt_plot[!,:mean] - dt_plot[!,:std]
dt_plot[!,:ymax] = dt_plot[!,:mean] + dt_plot[!,:std]
ggplot(data=dt_plot, aes(x=:period, y=:mean, colour=:generator)) +
    geom_line() + 
    geom_errorbar(aes(ymin=:ymin, ymax=:ymax), width=.2) + 
    geom_point() +
    facet_wrap(R".~prop") +
    theme_bw() |>
    p -> ggsave("www/german_dynamic_validity.png", plot = p);
load("www/german_dynamic_validity.png")
```

Interesting to see that there are some clear trends in the cost of recourse over time:

- what could be the reason for these dynamics?

```{julia}
dt_plot = groupby(results, [:period, :generator, :prop]) |>
    gdf -> combine(gdf, :cost .=> [mean, std] .=> [:mean, :std])
dt_plot[!,:ymin] = dt_plot[!,:mean] - dt_plot[!,:std]
dt_plot[!,:ymax] = dt_plot[!,:mean] + dt_plot[!,:std]
ggplot(data=dt_plot, aes(x=:period, y=:mean, colour=:generator)) +
    geom_line() + 
    geom_errorbar(aes(ymin=:ymin, ymax=:ymax), width=.2) + 
    geom_point() +
    facet_wrap(R".~prop") +
    theme_bw() |>
    p -> ggsave("www/german_dynamic_cost.png", plot = p);
load("www/german_dynamic_cost.png")
```

### Adjusting for imbalance

```{julia}
using MLDataUtils
X_train, y_train = oversample((transpose(X),y))
X_train = transpose(X_train);
model = bayes_logreg(X_train,y_train);
```

```{julia}
y_hat = predict(model,X)
rocplot(y,y_hat)
```