### 2022-04-13 paper/sections/introduction.qmd
more work on setting up submission

---
format: gfm
---

## Introduction {#intro}

...