### 2022-04-13 dev/submissions/aies/sections/cover.qmd
aies

## Front matter

### Author

- Full Name: {{< var author.name >}}
- Home page: {{< var author.url >}}
- Affiliation: {{< var author.affiliation >}}
- Affiliation URL: {{< var author.affiliation-url >}}
- Country of citizenship: {{< var author.nationality >}}

### PhD Progress

- Expected graduation date: {{< var phd.graduation-date >}}
- Years in program: {{< var phd.years >}}

### Research Advisor

- Full Name: {{< var advisor.name >}}

*Key words*: {{< var keywords >}}