### 2022-04-13 dev/submissions/aies/sections/support.qmd

### Letter of Support

Patrick has been my supervisee for ...