### 2022-04-25 dev/submissions/aies/sections/support.qmd

---
format:
    pdf: default
---

### Letter of Support

Patrick has been my supervisee for ...