### 2022-04-25 dev/submissions/aies/extended_abstract/extended_abstract.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{acmconf}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Dynamics in Algorithmic Recourse},
  pdfauthor={Patrick Altmeyer},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\ConferenceShortName{AIES '22}
\ConferenceName{Fifth AAAI/ACM Conference on Artificial Intelligence, Ethics, and Society}
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Dynamics in Algorithmic Recourse}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Trustworthy Artificial Intelligence for Finance and Economics}
\author{Patrick Altmeyer}
\date{April 25, 2022}

\begin{document}
\maketitle

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, sharp corners, interior hidden, enhanced, breakable, frame hidden]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Recent advances in artificial intelligence (AI) have propelled its
adoption in domains outside of computer science including health care,
bioinformatics and genetics. In finance, economics and other social
sciences, applications of AI are still relatively limited.
Decision-making in these fields has traditionally been guided by
Generalized Linear Models (GLM), which are theoretically founded,
interpretable and often sufficient to model relationships between
variables. Model interpretability is crucial in the social sciences
context, because inference is typically at least as important as
predictive performance. Decision-makers in the social sciences are also
typically required to explain their decisions to human stakeholders:
central bankers, for example, are held accountable by the public for the
policies they decide on. It is therefore not surprising that
practitioners and academics in these fields are reluctant to adopt AI
technologies that ultimately cannot be trusted. Deep learning models,
for example, are generally considered as black boxes and therefore
difficult to apply in a context that demands explanations. This PhD
project is focused on exploring and developing methodologies that
improve the trustworthiness of AI and thereby enable its application in
Finance and Economics.

The remainder of this extended abstract is structured as follows:
Section~\ref{sec-main} presents one of the research questions I have
investigated during the first months of my PhD: how do counterfactual
explanations handle dynamics? Section~\ref{sec-related} places this work
in the broader context of Trustworthy AI for Finance and Economics.

\hypertarget{sec-main}{%
\section{Dynamics in Algorithmic Recourse}\label{sec-main}}

\textbf{Counterfactual explanations} (CE) explain how inputs into a
model need to change for it to produce different outputs. They are
intuitive, simple and intrinsically linked to the potential outcome
framework for causal inference, which social scientists are familiar
with. Counterfactual explanations that involve realistic and actionable
changes can be used for the purpose of \textbf{Algorithmic Recourse}
(AR) to help individuals facing adverse decisions. An example relevant
to the Finance and Econonomics domain is consumer credit: in this
context AR can be used to guide individuals to improve their credit
worthiness, should they have previously been denied access to credit
based on a black-box decision-making system .

Existing work on CE and AR has largely been limited to the static
setting: given some classifier \(M: \mathcal{X} \mapsto \mathcal{Y}\) we
are interested in finding close (Wachter, Mittelstadt, and Russell
2017), actionable (Ustun, Spangher, and Liu 2019), plausible Schut et
al. (2021), sparse (Schut et al. 2021), diverse (Mothilal, Sharma, and
Tan 2020) and ideally causally founded counterfactual explanations
(Karimi, Schölkopf, and Valera 2021) for some individual \(x\). The
ability of counterfactual explanations to handle dynamics like data and
model shifts remains a largely unexplored research challenge at this
point (Verma, Dickerson, and Hines 2020). Only one recent work considers
the implications of \textbf{exogenous} domain shifts on the validity of
recourse (Upadhyay, Joshi, and Lakkaraju 2021). The authors propose a
simple minimax objective, that minimizes the counterfactual loss
function for a maximal domain and model shift. They show that their
approach yields more robust counterfactuals than existing approaches.

This project investigates \textbf{endogenous} domain and model shifts,
that is shifts that occur when AR is actually impemented by a proportion
of individuals and the classifier is updated in response.
Figure~\ref{fig-dynamics} illustrates this idea for a binary problem
involving a probabilistic classifier and a greedy counterfactual
generator proposed by Schut et al. (2021): AR leads to a domain shift,
which in turn causes a drastic model shift. As this game of implementing
AR and updating the classifier is repeated, individuals who receive and
implement algorithmic recourse end up forming a distinct subgroup inside
the target class, which may leave them vulnerable to discrimination.
Through future experiments we want to investigate if this phenomenon is
robust across different benchmark datasets and counterfactual
generators.

\begin{figure}

{\centering \includegraphics[width=2.60417in,height=\textheight]{www/dynamics.png}

}

\caption{\label{fig-dynamics}PLACEHOLDER: The dynamics of algorithmic
recourse.}

\end{figure}

\hypertarget{sec-related}{%
\section{Related and Future Work}\label{sec-related}}

\hypertarget{benchmarking-ce-in-julia}{%
\subsection{Benchmarking CE in Julia}\label{benchmarking-ce-in-julia}}

Alongside my research I have developed open-source implementations
related to explainable AI.
\href{https://www.paltmeyer.com/CounterfactualExplanations.jl/stable/}{CounterfactualExplanations.jl}
is a Julia package that can be used to generate counterfactual
explanations for models developed and trained not only in Julia, but
also in other popular programming languages like Python and R. I have
recently submitted the package along with a companion paper as a
proposal for a main talk at \href{https://juliacon.org/2022/}{JuliaCon}.
\href{https://www.paltmeyer.com/BayesLaplace.jl/dev/}{BayesLaplace.jl}
is a small Julia package that can be used to recover Bayesian
representations of deep neural networks through Laplace approximation in
a post-hoc manner. It is inspired by a recent paper (Daxberger et al.
2021) and has also been submitted to JuliaCon. Finally,
\href{https://github.com/pat-alt/deepvars}{deepvars} is an R package
that implements an approach towards vector autoregression that leverages
deep learning. This was originally my master's thesis and later
presented at the NeurIPS 2021 MLECON workshop. I have also published
several blog posts on explainable AI and probabilisitic ML in an effort
to make my research accessible to a broad audience.

\hypertarget{probabilistic-methods-for-realistic-counterfactual-explanations}{%
\subsection{Probabilistic methods for realistic counterfactual
explanations}\label{probabilistic-methods-for-realistic-counterfactual-explanations}}

Probabilistic machine learning can be leveraged in this context and more
generally facilitates inference and interpretability. It is also closely
related to Bayesian statistics, which has played an important role in
both finance and economics for many years.

To ensure that the generated explanations are realistic it is important
to understand which input-output pairs are likely and which are not. To
quantify their joint likelihood, previous work has either relied on
generative models or restricted the analysis to probabilistic models
that incorporate uncertainty in their predictions. While the former
approach is more versatile since it is applicable to both deterministic
and probabilistic models, the latter is computationally much more
efficient. In my work I want to explore how recent advances in post-hoc
uncertainty quantification can be leveraged to generate realistic and
unambiguous counterfactual explanations for any model.

\hypertarget{ce-for-time-series}{%
\subsection{CE for time series}\label{ce-for-time-series}}

Data sets in finance and economics typically involve time series data.
Therefore, I am naturally interested in the application of explainable
AI to sequential data, an area which has so far not been explored
extensively. In the future, I want to work on counterfactual
explanations for time series models. I am also interested in seeing if
and how Laplace approximation can be used for Bayesian deep learning
with time series data. I hope that the findings from both of these
projects can ultimately be used to build complex but interpretable time
series models for classification and forecasting in finance and
economics.

\hypertarget{explainable-black-box-models-for-time-series}{%
\subsection{Explainable black-box models for time
series}\label{explainable-black-box-models-for-time-series}}

For example, I would like to leverage effortless Bayesian deep learning
to make our proposed Deep Vector Autoregression model explainable.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-antoran2020getting}{}}%
Antorán, Javier, Umang Bhatt, Tameem Adel, Adrian Weller, and José
Miguel Hernández-Lobato. 2020. {``Getting a Clue: A Method for
Explaining Uncertainty Estimates.''} \emph{arXiv Preprint
arXiv:2006.06848}.

\leavevmode\vadjust pre{\hypertarget{ref-daxberger2021laplace}{}}%
Daxberger, Erik, Agustinus Kristiadi, Alexander Immer, Runa Eschenhagen,
Matthias Bauer, and Philipp Hennig. 2021. {``Laplace Redux-Effortless
Bayesian Deep Learning.''} \emph{Advances in Neural Information
Processing Systems} 34.

\leavevmode\vadjust pre{\hypertarget{ref-joshi2019towards}{}}%
Joshi, Shalmali, Oluwasanmi Koyejo, Warut Vijitbenjaronk, Been Kim, and
Joydeep Ghosh. 2019. {``Towards Realistic Individual Recourse and
Actionable Explanations in Black-Box Decision Making Systems.''}
\emph{arXiv Preprint arXiv:1907.09615}.

\leavevmode\vadjust pre{\hypertarget{ref-karimi2021algorithmic}{}}%
Karimi, Amir-Hossein, Bernhard Schölkopf, and Isabel Valera. 2021.
{``Algorithmic Recourse: From Counterfactual Explanations to
Interventions.''} In \emph{Proceedings of the 2021 ACM Conference on
Fairness, Accountability, and Transparency}, 353--62.

\leavevmode\vadjust pre{\hypertarget{ref-mothilal2020explaining}{}}%
Mothilal, Ramaravind K, Amit Sharma, and Chenhao Tan. 2020.
{``Explaining Machine Learning Classifiers Through Diverse
Counterfactual Explanations.''} In \emph{Proceedings of the 2020
Conference on Fairness, Accountability, and Transparency}, 607--17.

\leavevmode\vadjust pre{\hypertarget{ref-schut2021generating}{}}%
Schut, Lisa, Oscar Key, Rory Mc Grath, Luca Costabello, Bogdan
Sacaleanu, Yarin Gal, et al. 2021. {``Generating Interpretable
Counterfactual Explanations by Implicit Minimisation of Epistemic and
Aleatoric Uncertainties.''} In \emph{International Conference on
Artificial Intelligence and Statistics}, 1756--64. PMLR.

\leavevmode\vadjust pre{\hypertarget{ref-upadhyay2021towards}{}}%
Upadhyay, Sohini, Shalmali Joshi, and Himabindu Lakkaraju. 2021.
{``Towards Robust and Reliable Algorithmic Recourse.''} \emph{arXiv
Preprint arXiv:2102.13620}.

\leavevmode\vadjust pre{\hypertarget{ref-ustun2019actionable}{}}%
Ustun, Berk, Alexander Spangher, and Yang Liu. 2019. {``Actionable
Recourse in Linear Classification.''} In \emph{Proceedings of the
Conference on Fairness, Accountability, and Transparency}, 10--19.

\leavevmode\vadjust pre{\hypertarget{ref-verma2020counterfactual}{}}%
Verma, Sahil, John Dickerson, and Keegan Hines. 2020. {``Counterfactual
Explanations for Machine Learning: A Review.''} \emph{arXiv Preprint
arXiv:2010.10596}.

\leavevmode\vadjust pre{\hypertarget{ref-wachter2017counterfactual}{}}%
Wachter, Sandra, Brent Mittelstadt, and Chris Russell. 2017.
{``Counterfactual Explanations Without Opening the Black Box: Automated
Decisions and the GDPR.''} \emph{Harv. JL \& Tech.} 31: 841.

\end{CSLReferences}

\end{document}