### 2022-04-25 dev/submissions/aies/extended_abstract/extended_abstract.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{acmconf}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Dynamics in Algorithmic Recourse},
  pdfauthor={Patrick Altmeyer},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\ConferenceShortName{AIES '22}
\ConferenceName{Fifth AAAI/ACM Conference on Artificial Intelligence, Ethics, and Society}
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Dynamics in Algorithmic Recourse}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Trustworthy Artificial Intelligence for Finance and Economics}
\author{Patrick Altmeyer}
\date{April 25, 2022}

\begin{document}
\maketitle

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, breakable, sharp corners, interior hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, frame hidden]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Recent advances in artificial intelligence (AI) have propelled its
adoption in domains outside of computer science including health care,
bioinformatics and genetics. In finance, economics and other social
sciences, applications of AI are still relatively limited.
Decision-making in these fields has traditionally been guided by
Generalized Linear Models (GLM), which are theoretically founded,
interpretable and often sufficient to model relationships between
variables. Model interpretability is crucial in the social sciences
context, because inference is typically at least as important as
predictive performance. Decision-makers in the social sciences are also
typically required to explain their decisions to human stakeholders:
central bankers, for example, are held accountable by the public for the
policies they decide on. It is therefore not surprising that
practitioners and academics in these fields are reluctant to adopt AI
technologies that ultimately cannot be trusted. Deep learning models,
for example, are generally considered as black boxes and therefore
difficult to apply in a context that demands explanations. This PhD
project is focused on exploring and developing methodologies that
improve the trustworthiness of AI and thereby enable its application in
Finance and Economics.

The remainder of this extended abstract is structured as follows:
Section~\ref{sec-main} presents one of the research questions I have
investigated during the first months of my PhD: how do counterfactual
explanations handle dynamics? Section~\ref{sec-related} places this work
in the broader context of Trustworthy AI for Finance and Economics.

\hypertarget{sec-main}{%
\section{Dynamics in Algorithmic Recourse}\label{sec-main}}

\textbf{Counterfactual explanations} (CE) explain how inputs into a
model need to change for it to produce different outputs. They are
intuitive, simple and intrinsically linked to the potential outcome
framework for causal inference, which social scientists are familiar
with. Counterfactual explanations that involve realistic and actionable
changes can be used for the purpose of \textbf{Algorithmic Recourse}
(AR) to help individuals facing adverse decisions. An example relevant
to the Finance and Econonomics domain is consumer credit: in this
context AR can be used to guide individuals to improve their credit
worthiness, should they have previously been denied access to credit
based on a black-box decision-making system .

Existing work on CE and AR has largely been limited to the static
setting: given some classifier \(M: \mathcal{X} \mapsto \mathcal{Y}\) we
are interested in finding close (Wachter, Mittelstadt, and Russell
2017), actionable (Ustun, Spangher, and Liu 2019), plausible Schut et
al. (2021), sparse (Schut et al. 2021), diverse (Mothilal, Sharma, and
Tan 2020) and ideally causally founded counterfactual explanations
(Karimi, Schölkopf, and Valera 2021) for some individual \(x\). The
ability of counterfactual explanations to handle dynamics like data and
model shifts remains a largely unexplored research challenge at this
point (Verma, Dickerson, and Hines 2020). Only one recent work considers
the implications of \textbf{exogenous} domain shifts on the validity of
recourse (Upadhyay, Joshi, and Lakkaraju 2021). The authors propose a
simple minimax objective, that minimizes the counterfactual loss
function for a maximal domain and model shift. They show that their
approach yields more robust counterfactuals than existing approaches.

This project investigates \textbf{endogenous} domain and model shifts,
that is shifts that occur when AR is actually impemented by a proportion
of individuals and the classifier is updated in response.
Figure~\ref{fig-dynamics} illustrates this idea for a binary problem
involving a probabilistic classifier and a greedy counterfactual
generator proposed by Schut et al. (2021): AR leads to a domain shift,
which in turn causes a drastic model shift. As this game of implementing
AR and updating the classifier is repeated, the decision boundary moves
in the opposite direction of the original training samples in the target
class. We consider several aspects of these dynamics as problematic.
Firstly, as the decision boundary moves in the direction of the
non-target class, counterfactual paths become shorter: in the loan
example, individuals that previously would have been denied credit based
on their input features are suddenly considered as credit worthy.
Average default risk across all borrowers can therefore be expected to
increase. Conversely, lenders that anticipate or detect these sorts of
dynamics may choose to still deny credit to individuals that have
implemented AR, thereby comprimising the validity of AR.

To the best of my knowledge this is the first work investigating
endogenous dynamic in algorithmic recourse. In future experiments I want
to investigate how this phenomenon plays out across different benchmark
datasets (including German credit ({``German Credit Data''} 1994),
Boston Housing ({``Boston Housing Data''} 1993) and COMPAS (Larson et
al. 2016)). Furthermore, I want to asses to what extent the magnitude
and direction of domain and model shifts depends on the choice of the
counterfactual generator. To this end, I am currently supervising a
group of undergraduate students, who are tackling some of these tasks in
their final-year research project.

\begin{figure}

{\centering \includegraphics[width=2.60417in,height=\textheight]{www/dynamics.png}

}

\caption{\label{fig-dynamics}PLACEHOLDER: The dynamics of algorithmic
recourse.}

\end{figure}

\hypertarget{sec-related}{%
\section{Related and Future Work}\label{sec-related}}

\hypertarget{benchmarking-ce-in-julia}{%
\subsection{Benchmarking CE in Julia}\label{benchmarking-ce-in-julia}}

Until recently there existed only one open-source library that provides
a unifying approach to generate and benchmark counterfactual
explanations for models built and trained in Python (Pawelczyk et al.
2021). To address this limitation I have developed
\href{https://www.paltmeyer.com/CounterfactualExplanations.jl/stable/}{CounterfactualExplanations.jl}:
a Julia package that can be used to generate counterfactual explanations
for models developed and trained not only in Julia, but also in other
popular programming languages like Python and R. The package and
companion paper are currently pending acceptance for a main talk at
\href{https://juliacon.org/2022/}{JuliaCon}.

\hypertarget{probabilistic-methods-for-realistic-ce}{%
\subsection{Probabilistic Methods for Realistic
CE}\label{probabilistic-methods-for-realistic-ce}}

To ensure that the generated explanations are realistic it is important
to understand which input-output pairs are likely and which are not. To
quantify their joint likelihood, previous work has either relied on
generative models or restricted the analysis to probabilistic models
that incorporate uncertainty in their predictions. While the former
approach is more versatile since it is applicable to both deterministic
and probabilistic models, the latter is computationally much more
efficient. The approach proposed by Schut et al. (2021) and used to
generate the examples in Figure~\ref{fig-dynamics} falls into the latter
category. In future work I want to explore how recent advances in
post-hoc uncertainty quantification, most notably Laplace Redux
(Daxberger et al. 2021), can be leveraged to generate realistic and
unambiguous counterfactual explanations for any model.\footnote{For some
  initial work on this see my Julia implementation of Laplace Redux:
  \href{https://www.paltmeyer.com/BayesLaplace.jl/dev/}{BayesLaplace.jl}.}

\hypertarget{xai-for-time-series-models}{%
\subsection{XAI for time series
models}\label{xai-for-time-series-models}}

Datasets in Finance and Economics often involve autoregressive time
series, which is modelled differently than static, cross-sectional data.
One popular approach to macroeconomic forecasting, for example, is
Vector Autoregression (VAR), which involves an interpretable system of
linear equations. Recent work has shown that machine learning can be
leveraged in this context to improve forecasting performance, albeit at
the sacrifice of explainability (Altmeyer, Agusti, and Vidal-Quadras
Costa 2021).\footnote{For an open-source implementation of our approach
  see \href{https://github.com/pat-alt/deepvars}{deepvars}: an R package
  Deep Vector Autoregression}. Through my PhD research I hope to
ultimately adress that shortfall through explainable AI and
probabilistic methods.

\hypertarget{counterfactual-explanations}{%
\subsubsection{Counterfactual
Explanations}\label{counterfactual-explanations}}

I would like to investigate if and how CE can be used to leverage
machine learning for time series forecasting without sacrificing
explainability. In the context of conventional VAR counterfactual
outcomes are actually well-studied (Kilian and Lütkepohl 2017), which
makes macroeconomic forecasting and interesting case study. More
generally though, work on applying CE to black-box models for time
series data has so far been limited and therefore warrants attention.

\hypertarget{probabilistic-methods}{%
\subsubsection{Probabilistic Methods}\label{probabilistic-methods}}

I am also interested in seeing if and how Laplace approximation can be
used for Bayesian deep learning with time series data. I hope that the
findings from both of these projects can ultimately be used to build
complex but interpretable time series models for classification and
forecasting in finance and economics.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-altmeyer2021deep}{}}%
Altmeyer, Patrick, Marc Agusti, and Ignacio Vidal-Quadras Costa. 2021.
{``Deep Vector Autoregression for Macroeconomic Data.''}
\url{https://thevoice.bse.eu/wp-content/uploads/2021/07/ds21-project-agusti-et-al.pdf}.

\leavevmode\vadjust pre{\hypertarget{ref-antoran2020getting}{}}%
Antorán, Javier, Umang Bhatt, Tameem Adel, Adrian Weller, and José
Miguel Hernández-Lobato. 2020. {``Getting a Clue: A Method for
Explaining Uncertainty Estimates.''} \emph{arXiv Preprint
arXiv:2006.06848}.

\leavevmode\vadjust pre{\hypertarget{ref-boston1993}{}}%
{``Boston Housing Data.''} 1993.
\url{https://archive.ics.uci.edu/ml/machine-learning-databases/housing/}.

\leavevmode\vadjust pre{\hypertarget{ref-daxberger2021laplace}{}}%
Daxberger, Erik, Agustinus Kristiadi, Alexander Immer, Runa Eschenhagen,
Matthias Bauer, and Philipp Hennig. 2021. {``Laplace Redux-Effortless
Bayesian Deep Learning.''} \emph{Advances in Neural Information
Processing Systems} 34.

\leavevmode\vadjust pre{\hypertarget{ref-germancredit1994}{}}%
{``German Credit Data.''} 1994.
\url{https://archive.ics.uci.edu/ml/datasets/statlog+(german+credit+data)}.

\leavevmode\vadjust pre{\hypertarget{ref-joshi2019towards}{}}%
Joshi, Shalmali, Oluwasanmi Koyejo, Warut Vijitbenjaronk, Been Kim, and
Joydeep Ghosh. 2019. {``Towards Realistic Individual Recourse and
Actionable Explanations in Black-Box Decision Making Systems.''}
\emph{arXiv Preprint arXiv:1907.09615}.

\leavevmode\vadjust pre{\hypertarget{ref-karimi2021algorithmic}{}}%
Karimi, Amir-Hossein, Bernhard Schölkopf, and Isabel Valera. 2021.
{``Algorithmic Recourse: From Counterfactual Explanations to
Interventions.''} In \emph{Proceedings of the 2021 ACM Conference on
Fairness, Accountability, and Transparency}, 353--62.

\leavevmode\vadjust pre{\hypertarget{ref-kilian2017structural}{}}%
Kilian, Lutz, and Helmut Lütkepohl. 2017. \emph{Structural Vector
Autoregressive Analysis}. Cambridge University Press.

\leavevmode\vadjust pre{\hypertarget{ref-compas2016}{}}%
Larson, Jeff, Surya Mattu, Lauren Kirchner, and Julia Angwin. 2016.
{``COMPAS.''} \url{https://github.com/propublica/compas-analysis}.

\leavevmode\vadjust pre{\hypertarget{ref-mothilal2020explaining}{}}%
Mothilal, Ramaravind K, Amit Sharma, and Chenhao Tan. 2020.
{``Explaining Machine Learning Classifiers Through Diverse
Counterfactual Explanations.''} In \emph{Proceedings of the 2020
Conference on Fairness, Accountability, and Transparency}, 607--17.

\leavevmode\vadjust pre{\hypertarget{ref-pawelczyk2021carla}{}}%
Pawelczyk, Martin, Sascha Bielawski, Johannes van den Heuvel, Tobias
Richter, and Gjergji Kasneci. 2021. {``Carla: A Python Library to
Benchmark Algorithmic Recourse and Counterfactual Explanation
Algorithms.''} \emph{arXiv Preprint arXiv:2108.00783}.

\leavevmode\vadjust pre{\hypertarget{ref-schut2021generating}{}}%
Schut, Lisa, Oscar Key, Rory Mc Grath, Luca Costabello, Bogdan
Sacaleanu, Yarin Gal, et al. 2021. {``Generating Interpretable
Counterfactual Explanations by Implicit Minimisation of Epistemic and
Aleatoric Uncertainties.''} In \emph{International Conference on
Artificial Intelligence and Statistics}, 1756--64. PMLR.

\leavevmode\vadjust pre{\hypertarget{ref-upadhyay2021towards}{}}%
Upadhyay, Sohini, Shalmali Joshi, and Himabindu Lakkaraju. 2021.
{``Towards Robust and Reliable Algorithmic Recourse.''} \emph{arXiv
Preprint arXiv:2102.13620}.

\leavevmode\vadjust pre{\hypertarget{ref-ustun2019actionable}{}}%
Ustun, Berk, Alexander Spangher, and Yang Liu. 2019. {``Actionable
Recourse in Linear Classification.''} In \emph{Proceedings of the
Conference on Fairness, Accountability, and Transparency}, 10--19.

\leavevmode\vadjust pre{\hypertarget{ref-verma2020counterfactual}{}}%
Verma, Sahil, John Dickerson, and Keegan Hines. 2020. {``Counterfactual
Explanations for Machine Learning: A Review.''} \emph{arXiv Preprint
arXiv:2010.10596}.

\leavevmode\vadjust pre{\hypertarget{ref-wachter2017counterfactual}{}}%
Wachter, Sandra, Brent Mittelstadt, and Chris Russell. 2017.
{``Counterfactual Explanations Without Opening the Black Box: Automated
Decisions and the GDPR.''} \emph{Harv. JL \& Tech.} 31: 841.

\end{CSLReferences}

\end{document}