### 2022-04-29 dev/submissions/aies/sections/cover.qmd

---
format:
    pdf: default
---

## Personal details

|  | Author | Research Advisor |
|---------|:--------|:--------|
| Name     | {{< var author.name >}} | {{< var advisor.name >}} |  
| Email | {{< var author.email >}} | {{< var advisor.email >}} | 
| Phone | {{< var author.tel >}} | {{< var advisor.tel >}} | 
| Home page | {{< var author.url >}} | {{< var advisor.url >}} | 
| Affiliation | {{< var author.affiliation >}} | {{< var advisor.affiliation >}} | 
| Affiliation URL | {{< var author.affiliation-url >}} | {{< var advisor.affiliation-url >}} | 
| Citizenship | {{< var author.nationality >}} | {{< var advisor.nationality >}} | 
| Expected graduation date | {{< var phd.graduation-date >}} | --- | 
| Years in program | {{< var phd.years >}} | --- |

*Key words*: {{< var keywords >}}