### 2022-07-06 paper/sections/limitations.qmd

## Limitations and Future Work {#sec-limit}

The experimental setup proposed here is designed to mimic a real-world recourse process in a simple fashion. In practice, models are in fact updated on a regular basis [@upadhyay2021towards]. We also find it plausible to assume that the implementation of recourse happens periodically for different individuals, rather that all at once at time $t=0$. That being said, our experimental design is a vast over-simplification of potential real-world scenarios. In practice, any endogenous shifts that may occur can be expected to be entangled with exogenous shifts of the nature investigated in @upadhyay2021towards. We also make implicit assumptions about the utility functions of the involved agents that may well be too simple: individuals seeking recourse are assumed to always implement the proposed counterfactual explanations; conversely, the agent in charge of the model $M$ is assumed to always treat individuals that have implemented valid recourse as if they were truly now in the target class. Relating this back to the consumer credit example, we assume that the would-be borrowers are always willing and able to implement recourse and the bank is always willing to provide credit as would-be borrowers move across the decision boundary. In practice it is doubtful that agents behave according to such simple rules.

Nonetheless, we think that our simple framework offers a starting point for future work on recourse dynamics (both endogenous and exogenous dynamics).