### 2022-07-26 dev/notebooks/mnist.qmd

---
title: Experiments with Credit Data
jupyter: julia-1.7
---

```{julia}
using Pkg; Pkg.activate("dev")
```

```{julia}
include("dev/utils.jl")
using AlgorithmicRecourseDynamics
using CounterfactualExplanations, Flux, Plots, PlotThemes, Random, LinearAlgebra
theme(:wong)
output_path = output_dir("mnist")
www_path = www_dir("mnist")
```

```{julia}
using Flux
using CounterfactualExplanations.Data: mnist_data, mnist_model, mnist_ensemble, mnist_vae
X, ys = mnist_data()
model = mnist_model()
ensemble = mnist_ensemble()
vae = mnist_vae()
```

```{julia}
n_per_class = 100
test_ratio = 0.2
train_data = CounterfactualData(X,ys)
using AlgorithmicRecourseDynamics.Models: undersample
train_data = undersample(train_data, n_per_class)
train_data.generative_model = vae
using MLDatasets
test_data = MNIST(:test)
X_test, y_test = test_data[:]
X_test = Flux.flatten(X_test)
y_test = Flux.onehotbatch(y_test,0:9)
test_data = undersample(CounterfactualData(X_test, y_test),Int(round(n_per_class*test_ratio)))
```

```{julia}
M = FluxModel(model, :classification_multi)
M_ens = FluxEnsemble(ensemble, :classification_multi)
models = Dict(:FluxModel => M, :FluxEnsemble => M_ens)
generators = Dict(
    :Greedy=>GreedyGenerator(), 
    :Generic=>GenericGenerator(),
    :REVISE=>REVISEGenerator(),
    # :DICE=>DiCEGenerator(),
)
```

```{julia}
experiment = set_up_experiment(train_data,test_data,models,generators;target=5)
```

```{julia}
n_evaluations = 1
n_rounds = 5
evaluate_every = Int(round(n_rounds/n_evaluations))
using Serialization
results = run_experiment(experiment;save_path=output_path,evaluate_every=evaluate_every,n_rounds=n_rounds)
Serialization.serialize(joinpath(output_path,"results.bson"),results)
```