### 2022-07-28 dev/notebooks/housing.qmd
implemented idea about gravitational generator

---
title: Experiments with Housing Data
jupyter: julia-1.7
---

```{julia}
using Pkg; Pkg.activate("dev")
```

```{julia}
include("dev/utils.jl")
using AlgorithmicRecourseDynamics
using CounterfactualExplanations, Flux, Plots, PlotThemes, Random, LaplaceRedux, LinearAlgebra
theme(:wong)
output_path = output_dir("real_world")
www_path = www_dir("real_world")
```

## Training the classifiers

```{julia}
using MLDatasets, Statistics, StatsBase
X = BostonHousing.features()
y = BostonHousing.targets()
y = Float64.(y .>= median(y)); # binary target
dt = fit(ZScoreTransform, X, dims=2)
StatsBase.transform!(dt, X)
data = CounterfactualData(X,y)
```

```{julia}
using CounterfactualExplanations.DataPreprocessing: unpack
bs = 50
function data_loader(data::CounterfactualData)
    X, y = unpack(data)
    data = Flux.DataLoader((X,y),batchsize=bs)
    return data
end
model_params = (batch_norm=true,n_hidden=128,dropout=true,p_dropout=0.1)
models = [:FluxModel, :FluxEnsemble]
models = Dict([(model,getfield(AlgorithmicRecourseDynamics.Models, model)(data; model_params...)) for model in models])
```

```{julia}
using AlgorithmicRecourseDynamics.Models
# Data:
data_train, data_test = Models.train_test_split(data)
n_epochs=100
map!(model -> Models.train(model, data_train; n_epochs=n_epochs, data_loader=data_loader), values(models))

```

```{julia}

generators = Dict(
    :Greedy=>GreedyGenerator(), 
    :Generic=>GenericGenerator(),
    :REVISE=>REVISEGenerator(),
    :DICE=>DiCEGenerator()
)
```

```{julia}
# Prepare data and model:
using Random
Random.seed!(1234)
using StatsBase
dt = fit(ZScoreTransform, X, dims=2)
StatsBase.transform!(dt, X)
xs = Flux.unstack(X,2)
data = zip(xs,y)
nn = Models.build_model(input_dim=size(X)[1], n_hidden=100)
loss(x, y) = Flux.Losses.logitbinarycrossentropy(nn(x), y)
```

The model achieves decent training accuracy

```{julia}
run = false
if run
  # Train model:
  using Flux.Optimise: update!, ADAM
  using Statistics, StatsBase
  opt = ADAM()
  epochs = 100
  avg_loss(data) = mean(map(d -> loss(d[1],d[2]), data))
  accuracy(data) = sum(map(d ->round.(Flux.σ.(nn(d[1]))) .== d[2], data))[1]/length(data)

  using Plots
  anim = Animation()
  avg_l = [avg_loss(data)]
  p1 = scatter( ylim=(0,avg_l[1]), xlim=(0,epochs), legend=false, xlab="Epoch", title="Average loss")
  acc = [accuracy(data)]
  p2 = scatter( ylim=(0.5,1), xlim=(0,epochs), legend=false, xlab="Epoch", title="Accuracy")

  τ = 1.0
  stopping_criterium_reached = accuracy(data) >= τ
  epoch = 1

  while epoch <= epochs && !stopping_criterium_reached
    for d in data
      gs = gradient(params(nn)) do
        l = loss(d...)
      end
      update!(opt, params(nn), gs)
    end
    avg_l = vcat(avg_l,avg_loss(data))
    plot!(p1, [0:epoch], avg_l, color=1)
    scatter!(p1, [0:epoch], avg_l, color=1)
    acc = vcat(acc,accuracy(data))
    plot!(p2, [0:epoch], acc, color=1)
    scatter!(p2, [0:epoch], acc, color=1)
    plt=plot(p1,p2, size=(600,300))
    frame(anim, plt)

    # Check if desired accuracy reached:
    stopping_criterium_reached = accuracy(data) >= τ

    epoch += 1

  end

  gif(anim, joinpath(www_folder, "single_nn.gif"), fps=10)

  using BSON: @save
  @save joinpath(output_folder, "nn.bson") nn

end
```

![](www/boston_housing/single_nn.gif)

Next we will build and train a deep ensemble.

```{julia}
opt = ADAM()
loss_type = :logitbinarycrossentropy
run = false
if run
    K = 50
    ensemble = Models.build_ensemble(K,kw=(input_dim=size(X)[1], n_hidden=100));
    ensemble, anim = Models.forward(ensemble, data, opt, n_epochs=30, plot_every=10, loss_type=loss_type); # fit the ensemble
    Models.save_ensemble(ensemble, root=joinpath(output_folder, "ensemble")) # save to disk
    gif(anim, joinpath(www_folder, "ensemble_loss.gif"), fps=25);
end
```

![](www/boston_housing/ensemble_loss.gif)