### 2022-08-08 dev/notebooks/real_world.qmd

---
title: Experiments with Credit Data
jupyter: julia-1.7
---

```{julia}
using Pkg; Pkg.activate("dev")
```

```{julia}
include("dev/utils.jl")
using AlgorithmicRecourseDynamics
using CounterfactualExplanations, Flux, Plots, PlotThemes, Random, LaplaceRedux, LinearAlgebra
theme(:wong)
output_path = output_dir("real_world")
www_path = www_dir("real_world")
```

```{julia}
max_obs = 10000
data_sets = AlgorithmicRecourseDynamics.Data.load_real_world(max_obs)
```

```{julia}
using CounterfactualExplanations.DataPreprocessing: unpack
bs = 50
function data_loader(data::CounterfactualData)
    X, y = unpack(data)
    data = Flux.DataLoader((X,y),batchsize=bs)
    return data
end
model_params = (batch_norm=false,n_hidden=32,n_layers=3,dropout=true,p_dropout=0.25)
```

```{julia}
models = [
    :LogisticRegression, 
    :FluxModel, 
    :FluxEnsemble
]
generators = Dict(
    :Greedy=>GreedyGenerator(), 
    :Generic=>GenericGenerator(),
    :REVISE=>REVISEGenerator(),
    :DICE=>DiCEGenerator(),
    :Gravitational=>GravitationalGenerator(λ=[0.1,2]),
)
```

```{julia}
experiments = set_up_experiments(
    data_sets,models,generators; 
    pre_train_models=100, model_params=model_params, 
    data_loader=data_loader
)
```

```{julia}
# #| eval: false
# using AlgorithmicRecourseDynamics.Models: model_evaluation
# plts = []
# for (exp_name, exp_) in experiments
#     for (M_name, M) in exp_.models
#         score = round(model_evaluation(M, exp_.test_data),digits=2)
#         plt = plot(M, exp_.test_data; title="$exp_name;\n $M_name ($score)", dim_red=:umap)
#         # # Errors:
#         # ids = findall(vec(round.(probs(M, exp_.test_data.X)) .!= exp_.test_data.y))
#         # x_wrongly_labelled = exp_.test_data.X[:,ids]
#         # scatter!(plt, x_wrongly_labelled[1,:], x_wrongly_labelled[2,:], ms=7.5, color=:red, label="")
#         plts = vcat(plts..., plt)
#     end
# end
# plt = plot(plts..., layout=(length(choices),length(models)),size=(length(choices)*300,length(models)*300))
# savefig(plt, joinpath(www_path,"models_test_before.png"))
```

```{julia}
# #| eval: false
# using AlgorithmicRecourseDynamics.Models: model_evaluation
# plts = []
# for (exp_name, exp_) in experiments
#     for (M_name, M) in exp_.models
#         score = round(model_evaluation(M, exp_.train_data),digits=2)
#         plt = plot(M, exp_.train_data; title="$exp_name;\n $M_name ($score)", dim_red=:umap)
#         # # Errors:
#         # ids = findall(vec(round.(probs(M, exp_.train_data.X)) .!= exp_.train_data.y))
#         # x_wrongly_labelled = exp_.train_data.X[:,ids]
#         # scatter!(plt, x_wrongly_labelled[1,:], x_wrongly_labelled[2,:], ms=7.5, color=:red, label="")
#         plts = vcat(plts..., plt)
#     end
# end
# plt = plot(plts..., layout=(length(choices),length(models)),size=(length(choices)*300,length(models)*300))
# savefig(plt, joinpath(www_path,"models_train_before.png"))
```

## Strict convergence

```{julia}
n_evals = 5
n_rounds = 50
evaluate_every = Int(round(n_rounds/n_evals))
n_folds = 5
n_bootstrap = 1
T = 500
using Serialization
results = run_experiments(
    experiments;
    save_path=output_path,evaluate_every=evaluate_every,n_rounds=n_rounds, n_folds=n_folds, n_bootstrap=n_bootstrap, T=T,
    convergence=:strict, save_name_suffix="strict"
)
Serialization.serialize(joinpath(output_path,"results_strict.jls"),results)
```

## Simple convergence

```{julia}
n_evals = 5
n_rounds = 50
evaluate_every = Int(round(n_rounds/n_evals))
n_folds = 5
n_bootstrap = 1
using Serialization
results = run_experiments(
    experiments;save_path=output_path,evaluate_every=evaluate_every,n_rounds=n_rounds, n_folds=n_folds, n_bootstrap=n_bootstrap
)
Serialization.serialize(joinpath(output_path,"results.jls"),results)
```

```{julia}
# using AlgorithmicRecourseDynamics.Models: model_evaluation
# plot_dict = Dict(key => Dict() for (key,val) in results)
# fold = 1
# for (name, res) in results
#     exp_ = res.experiment
#     plot_dict[name] = Dict(key => [] for (key,val) in exp_.generators)
#     rec_sys = exp_.recourse_systems[fold]
#     sys_ids = collect(exp_.system_identifiers)
#     M = length(rec_sys)
#     for m in 1:M
#         model_name, generator_name = sys_ids[m]
#         M = rec_sys[m].model
#         score = round(model_evaluation(M, exp_.test_data),digits=2)
#         plt = plot(M, exp_.test_data; title="$name;\n $model_name ($score)", dim_red=:umap)
#         # # Errors:
#         # ids = findall(vec(round.(probs(M, exp_.test_data.X)) .!= exp_.test_data.y))
#         # x_wrongly_labelled = exp_.test_data.X[:,ids]
#         # scatter!(plt, x_wrongly_labelled[1,:], x_wrongly_labelled[2,:], ms=7.5, color=:red, label="")
#         plot_dict[name][generator_name] = vcat(plot_dict[name][generator_name], plt)
#     end
# end
# plot_dict = Dict(key => reduce(vcat, [plots[key] for plots in values(plot_dict)]) for (key, value) in generators)
# for (name, plts) in plot_dict
#     plt = plot(plts..., layout=(length(choices),length(models)),size=(length(choices)*300,length(models)*300))
#     savefig(plt, joinpath(www_path,"models_test_after_$name.png"))
# end
```

```{julia}
# using AlgorithmicRecourseDynamics.Models: model_evaluation
# plot_dict = Dict(key => Dict() for (key,val) in results)
# fold = 1
# for (name, res) in results
#     exp_ = res.experiment
#     plot_dict[name] = Dict(key => [] for (key,val) in exp_.generators)
#     rec_sys = exp_.recourse_systems[fold]
#     sys_ids = collect(exp_.system_identifiers)
#     M = length(rec_sys)
#     for m in 1:M
#         model_name, generator_name = sys_ids[m]
#         M = rec_sys[m].model
#         data = rec_sys[m].data
#         score = round(model_evaluation(M, data),digits=2)
#         plt = plot(M, data; title="$name;\n $model_name ($score)", dim_red=:umap)
#         # # Errors:
#         # ids = findall(vec(round.(probs(M, data.X)) .!= data.y))
#         # x_wrongly_labelled = data.X[:,ids]
#         # scatter!(plt, x_wrongly_labelled[1,:], x_wrongly_labelled[2,:], ms=7.5, color=:red, label="")
#         plot_dict[name][generator_name] = vcat(plot_dict[name][generator_name], plt)
#     end
# end
# plot_dict = Dict(key => reduce(vcat, [plots[key] for plots in values(plot_dict)]) for (key, value) in generators)
# for (name, plts) in plot_dict
#     plt = plot(plts..., layout=(length(choices),length(models)),size=(length(choices)*300,length(models)*300))
#     savefig(plt, joinpath(www_path,"models_train_after_$name.png"))
# end
```

```{julia}
using Serialization
results = Serialization.deserialize(joinpath(output_path,"results.jls"))
```

### Plots

```{julia}
# plot_dict = Dict()
# for (data_name, res) in results
#     output_data = res.output
#     metrics = unique(output_data.name)
#     plots = Dict(metric => plot(res, metric) for metric in metrics)
#     plot_dict[data_name] = plots
# end
```

```{julia}
using AlgorithmicRecourseDynamics.Models: model_evaluation
plot_dict = Dict(key => Dict() for (key,val) in results)
fold = 1
for (name, res) in results
    exp_ = res.experiment
    plot_dict[name] = Dict(key => [] for (key,val) in exp_.generators)
    rec_sys = exp_.recourse_systems[fold]
    sys_ids = collect(exp_.system_identifiers)
    M = length(rec_sys)
    for m in 1:M
        model_name, generator_name = sys_ids[m]
        M = rec_sys[m].model
        score = round(model_evaluation(M, exp_.test_data),digits=2)
        plt = plot(M, exp_.test_data; title="$name;\n $model_name ($score)", dim_red=:umap)
        # # Errors:
        # ids = findall(vec(round.(probs(M, exp_.test_data.X)) .!= exp_.test_data.y))
        # x_wrongly_labelled = exp_.test_data.X[:,ids]
        # scatter!(plt, x_wrongly_labelled[1,:], x_wrongly_labelled[2,:], ms=7.5, color=:red, label="")
        plot_dict[name][generator_name] = vcat(plot_dict[name][generator_name], plt)
    end
end
plot_dict = Dict(key => reduce(vcat, [plots[key] for plots in values(plot_dict)]) for (key, value) in generators)
for (name, plts) in plot_dict
    plt = plot(plts..., layout=(length(choices),length(models)),size=(length(choices)*300,length(models)*300))
    savefig(plt, joinpath(www_path,"models_test_after_$(name)_strict.png"))
end
```

```{julia}
using AlgorithmicRecourseDynamics.Models: model_evaluation
plot_dict = Dict(key => Dict() for (key,val) in results)
fold = 1
for (name, res) in results
    exp_ = res.experiment
    plot_dict[name] = Dict(key => [] for (key,val) in exp_.generators)
    rec_sys = exp_.recourse_systems[fold]
    sys_ids = collect(exp_.system_identifiers)
    M = length(rec_sys)
    for m in 1:M
        model_name, generator_name = sys_ids[m]
        M = rec_sys[m].model
        data = rec_sys[m].data
        score = round(model_evaluation(M, data),digits=2)
        plt = plot(M, data; title="$name;\n $model_name ($score)", dim_red=:umap)
        # # Errors:
        # ids = findall(vec(round.(probs(M, data.X)) .!= data.y))
        # x_wrongly_labelled = data.X[:,ids]
        # scatter!(plt, x_wrongly_labelled[1,:], x_wrongly_labelled[2,:], ms=7.5, color=:red, label="")
        plot_dict[name][generator_name] = vcat(plot_dict[name][generator_name], plt)
    end
end
plot_dict = Dict(key => reduce(vcat, [plots[key] for plots in values(plot_dict)]) for (key, value) in generators)
for (name, plts) in plot_dict
    plt = plot(plts..., layout=(length(choices),length(models)),size=(length(choices)*300,length(models)*300))
    savefig(plt, joinpath(www_path,"models_train_after_$(name)_strict.png"))
end
```

```{julia}
using Serialization
results = Serialization.deserialize(joinpath(output_path,"results_strict.jls"))
```