### 2022-08-08 dev/paper/sections/empirical.qmd

# Experiments {#sec-empirical}

1. Shift of focus from individual to group of individuals (related: https://www.researchgate.net/publication/353073138_Generating_Collective_Counterfactual_Explanations_in_Score-Based_Classification_via_Mathematical_Optimization)
2. Convergence criterium matters: terminating once threshold probability is reached may not be optimal (see e.g. REVISE)
3. Optimizer choice matters: dimensionality is typically low, so no obvious benefit to using ADAM.
    - This might be better placed in JuliaCon proceedings, perhaps backed by small blog post on the matter. 
4. Mitigating strategy: penaliye distance from centroid.