### 2022-08-09 dev/paper/sections/introduction.qmd
some work on rephrasing narrative in intro

# Introduction {#sec-intro}

Recent advances in Artificial Intelligence (AI) have propelled its adoption in scientific domains outside of Computer Science including Healthcare, Bioinformatics, Genetics and the Social Sciences. While this has in many cases brought benefits in terms of efficiency, state-of-the-art models like Deep Neural Networks (DNN) have also given rise a new type of principal-agent problem in the context of data-driven decision-making. It involves a group of **principals** - i.e. human stakeholders - that fail to understand the behaviour of their **agent** - i.e. the model used for automated decision-making [@borch2022machine].

Models or algorithms that fall into this category are typically referred to **black-box** models. Despite their shortcomings, black-box models have grown in popularity in recent years and have at times created undesirable societal outcomes [@o2016weapons]. The scientific community has tackled this issue from two different angles: while some have appealed for a strict focus on inherently iterpretable models [@rudin2019stop], others have investigated different ways to explain the behaviour of black-box models. These two sub-domains can be broadly referred to as **interpretable AI** and **explainable AI** (XAI), respectively.

Among the approaches to XAI that have recently grown in popularity are **Counterfactual Explanations** (CE). They explain how inputs into a model need to change for it to produce different outputs. Counterfactual Explanations that involve realistic and actionable changes can be used for the purpose of **Algorithmic Recourse** (AR) to help individuals who face adverse outcomes. An example relevant to the Social Sciences is consumer credit: in this context AR can be used to guide individuals in improving their creditworthiness, should they have previously been denied access to credit based on an automated decision-making system. A meaningful recourse recommendation for a denied applicant could be: *"If your net savings rate had been 10% of your monthly income instead of the actual 8%, your application would have been successful. See if you can temporarily cut down on consumption."* In the remainder of this paper we will use both terminologies - recourse and counterfactual - interchangeably to refer to situations where counterfactuals are generated with the intent to provide individual recourse.

Existing work in this field has largely worked in a static setting: various approaches have been proposed to generate counterfactuals for a given individual that is subject to some pre-trained model. More recent work has compared different approaches within this static setting [@pawelczyk2021carla]. In this work we go one step further and ask ourselves: what happens if recourse is provided and implemented repeatedly? What types of dynamics are introduced and how do different counterfactual generators compare in this context?

Research on algorithmic recourse has also so far typically addressed the issue from the perspective of one single individual and has indeed been referred to as **individual recourse** in some places. Arguably though, most real-world applications that warrant algorithmic recourse involve potentially large groups of individuals typically competing for scarce resources. Our work demonstrates that in such scenarios, choices made by or for one single individual are likely to affect the broader collective of individuals in ways that current approaches to AR fail to account for. More specifically, we argue that a strict focus on minimizing the private costs faced by individuals may be too narrow an objective.

@fig-poc illustrates this idea for a binary problem involving a probabilistic classifier and the counterfactual generator proposed by @wachter2017counterfactual: the implementation of AR for a subset of individuals leads to a domain shift (b), which in turn triggers a model shift (c). As this game of implementing AR and updating the classifier is repeated, the decision boundary moves away from training samples that were originally in the target class (d). We refer to these types of dynamics as **endogenous** because they are induced by the implementation of recourse itself. The term **macrodynamics** is borrowed from the economics literature and used to describe processes involving whole groups or societies.

![Dynamics in Algorithmic Recourse: we have a simple Bayesian model trained for binary classification (a); the implementation of AR for a random subset of individuals leads to a domain shift (b); as the classifier is retrained we observe a model shift (c); as this process is repeated, the decision boundary moves away from the target class (d).](www/poc.png){#fig-poc fig.pos="h" width=45%}

We think that these types of endogenous dynamics may be problematic and warrant our attention. Firstly, model shifts may inadvertently change classification outcomes for individuals who never received and implemented recourse. Secondly and relatedly, we observe in @fig-poc that as the decision boundary moves in the direction of the non-target class, counterfactual paths become shorter: in the consumer credit example, individuals that previously would have been denied credit based on their input features are suddenly considered as creditworthy. Average default risk across all borrowers can therefore be expected to increase. Conversely, lenders that anticipate such dynamics may choose to refrain from offering recourse (and hence credit) to more than just a tiny share of individuals. In that latter and perhaps more likely scenario, the probability of being offered recourse decreases with every individual that implements recourse: in other words, the actions of first-movers exert a negative externality on future would-be borrowers.

To the best of our knowledge this is the first work investigating endogenous macrodynamics in AR. Our contributions to the state of knowledge are as follows: firstly, we posit a compelling argument that calls for a novel perspective on algorithmic recourse extending our focus from single individuals to groups. Secondly, we introduce an experimental framework extending previous work by @altmeyer2022CounterfactualExplanations, which enables us to study macrodynamics of algorithmic recourse through simulations. To this end we propose a number of novel evaluation metrics that can be used to qunatify and benchmark the macrodynmics introduced by different counterfactual generators. Thirdly, we use this framework to provide a first in-depth analysis of endogenous recourse dynamics induced by various popular counterfactual generators including @wachter2017counterfactual, @schut2021generating, @joshi2019towards, @mothilal2020explaining and @antoran2020getting. Finally, we discuss what drives endogenous dynamics and strategies to mitigate them.

The remainder of the paper is structured as follows: @sec-related places our work in the broader context of related literature. @sec-method presents our methodology and data. @sec-empirical presents our empirical findings which are then discussed in the broader context of the literature in @sec-discussion. We also point to some of the limitations or our work as well as avenues for future research in @sec-limit. Finally, @sec-conclusion concludes.