### 2022-08-09 dev/notebooks/data_preprocessing.qmd/synthetic_data.qmd
new and larger synthetic datasets

---
title: Generating Synthetic Data
jupyter: julia-1.7
---

```{julia}
using Pkg; Pkg.activate("dev")
```

```{julia}
include("dev/utils.jl")
using AlgorithmicRecourseDynamics
theme(:wong)
output_path = output_dir("synthetic")
www_path = www_dir("synthetic")
data_path = data_dir("synthetic")
```

```{julia}
using MLJ, DataFrames, CSV, VegaLite
n = 1000
p = 2

using Random
Random.seed!(42)

# Linearly separable:
X, y = make_blobs(n, p; centers=2, center_box=(-2 => 2), cluster_std=0.1)
df = DataFrame(X)
df.target = y
df |> @vlplot(:point, x=:x1, y=:x2, color = :"target:n")
CSV.write(joinpath(data_path, "linearly_separable.csv"),df)

# Overlapping:
X, y = make_blobs(n, p; centers=2, center_box=(-2 => 2), cluster_std=0.5)
df = DataFrame(X)
df.target = y
df |> @vlplot(:point, x=:x1, y=:x2, color = :"target:n")
CSV.write(joinpath(data_path, "overlapping.csv"),df)

# Circles:
X, y = make_circles(n; noise=0.15, factor=0.01)
df = DataFrame(X)
df.target = y
df |> @vlplot(:point, x=:x1, y=:x2, color = :"target:n")
CSV.write(joinpath(data_path, "circles.csv"),df)

# Moon:
X, y = make_moons(n)
df = DataFrame(X)
df.target = y
df |> @vlplot(:point, x=:x1, y=:x2, color = :"target:n")
CSV.write(joinpath(data_path, "moons.csv"),df)
```