### 2022-08-10 dev/notebooks/endo_roar_generator.qmd
added another generator that simply penalizes the loss the model incurs on the counterfactual.

---
title: EndoROAR Generator
jupyter: julia-1.7
---

```{julia}
using Pkg; Pkg.activate("dev")
```

```{julia}
include("dev/utils.jl")
using AlgorithmicRecourseDynamics
using CounterfactualExplanations, Flux, Plots, PlotThemes, Random, LaplaceRedux, LinearAlgebra
theme(:wong)
output_path = output_dir("generator")
www_path = www_dir("generator")
```

# `EndoROARGenerator`

```{julia}
using MLJ
N = 1000
X, ys = make_blobs(N, 2; centers=2, as_table=false, center_box=(-5 => 5), cluster_std=0.1)
ys .= ys.==2
X = X'
xs = Flux.unstack(X,2)
data = zip(xs,ys)
counterfactual_data = CounterfactualData(X,ys')
```

```{julia}
using Flux
nn = Chain(Dense(2,1))
using Flux.Optimise: update!, ADAM
opt = ADAM()
epochs = 100
loss(x, y) = Flux.Losses.logitbinarycrossentropy(nn(x), y)
avg_loss(data) = mean(map(d -> loss(d[1],d[2]), data))
show_every = epochs/10
grad_norms = []
using LinearAlgebra
for epoch = 1:epochs
    grads_ = []
    for d in data
        gs = gradient(Flux.params(nn)) do
            l = loss(d...)
        end
        update!(opt, Flux.params(nn), gs)
        grads_ = vcat(grads_..., gs)
    end
    if epoch % show_every == 0
        println("Epoch " * string(epoch))
        @show avg_loss(data)
    end
    grad_norms = vcat(grad_norms..., norm(grads_))
end
```

```{julia}
M = FluxModel(nn)
```

```{julia}
x = select_factual(counterfactual_data, rand(1:size(X)[2])) 
y = round(probs(M, x)[1])
target = ifelse(y==1.0,0.0,1.0) # opposite label as target
T = 100
```

```{julia}
generator = GenericGenerator()
ce = generate_counterfactual(x, target,counterfactual_data, M, generator)
```

```{julia}
x_ = path(ce)
y_ = CounterfactualExplanations.Counterfactuals.counterfactual_label_path(ce)
data_ = zip(x_,y_)
grad_norms = []
loss_ = []
for d in data_
    gs = gradient(Flux.params(nn)) do
        loss(d...)
    end
    loss_ = vcat(loss_...,loss(d...))
    grad_norms = vcat(grad_norms..., norm(gs))
end
```

```{julia}
Λ₂ = [0.1, 1, 5]
counterfactuals_strict = []
generators = []
for λ₂ ∈ Λ₂  
    λ = [0.1, λ₂]
    generator = EndoROARGenerator(λ=λ)
    generators = vcat(generators..., generator)
    counterfactuals_strict = vcat(
      counterfactuals_strict...,
      generate_counterfactual(x, target, counterfactual_data, M, generator; convergence=:strict, T=T)
    )
end
```

```{julia}
theme(:wong)
T_ = 500
plts = []
for i ∈ 1:length(Λ₂)
    λ₂ = Λ₂[i]
    counterfactual = counterfactuals_strict[i]  
    plt = plot(counterfactual, plot_up_to=minimum([T,T_]), title="λ₂=$(λ₂)")
    plts = vcat(plts..., plt)
end
plt = plot(plts..., size=(1200,300), layout=(1,3))
# savefig(plt, joinpath(www_path,"endo_roar_generator.png"))
```

```{julia}
generators = Dict(
    :Generic => GenericGenerator(),
    :ROAR => EndoROARGenerator()
)
counterfactuals = Dict([name => generate_counterfactual(x, target, counterfactual_data, M, gen; convergence=:strict, T=T) for (name, gen) in generators])
```

```{julia}
plts = []
for (name,ce) ∈ counterfactuals
    plt = plot(ce, plot_up_to=minimum([T,T_]), title=name)
    plts = vcat(plts..., plt)
end
plt = plot(plts..., size=(800,300), layout=(1,2))
```