### 2022-08-11 dev/paper/sections/empirical_2.qmd

# Experiments {#sec-empirical-2}

## Endogenous Macrodynamics

## Potential Mitigation Strategies