### 2022-08-12 dev/paper/paper.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  conference]{IEEEtran}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithm}% http://ctan.org/pkg/algorithms
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{placeins}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Endogenous Macrodynamics in Algorithmic Recourse},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Endogenous Macrodynamics in Algorithmic Recourse}
\author{}
\date{}

\begin{document}
\maketitle
\begin{abstract}
Existing work on Counterfactual Explanations (CE) and Algorithmic
Recourse (AR) has largely been limited to the static setting and focused
on single individuals: given some estimated model the goal is to find
valid counterfactuals for individual instance that fulfill various
desiderata. The ability of such counterfactuals to handle dynamics like
data and model drift remains a largely unexplored research challenge at
this point. There has also been surprisingly little work on the related
question of how the actual implementation of recourse by one individual
may affect other individuals. Through this work we aim to close that gap
by systematizing and extending existing knowledge. We first show that
many of the existing methodologies can be collectively described by a
generalized framework. We then argue that the existing framework fails
to account for a hidden external cost of recourse, that only reveals
itself when studying the endogenous dynamics of recourse at the group
level. Through simulation experiments involving various popular
counterfactual generators and several benchmark datasets, we generate a
total XX million Counterfactual Explanations and study the resulting
domain and model shifts. We find that the induced shifts are substantial
enough to likely impede the applicability Algorithmic Recourse in
practice. Fortunately, we find various potential mitigation strategies
that can be used in combination with existing approaches. Our simulation
framework for studying recourse dynamics is fast and open-sourced.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, enhanced, frame hidden, breakable, interior hidden]}{\end{tcolorbox}}\fi

\hypertarget{sec-intro}{%
\section{Introduction}\label{sec-intro}}

Recent advances in Artificial Intelligence (AI) have propelled its
adoption in scientific domains outside of Computer Science including
Healthcare, Bioinformatics, Genetics and the Social Sciences. While this
has in many cases brought benefits in terms of efficiency,
state-of-the-art models like Deep Neural Networks (DNN) have also given
rise a new type of principal-agent problem in the context of data-driven
decision-making. It involves a group of \textbf{principals} - i.e.~human
stakeholders - that fail to understand the behaviour of their
\textbf{agent} - i.e.~the model used for automated decision-making
\protect\hyperlink{ref-borch2022machine}{{[}1{]}}.

Models or algorithms that fall into this category are typically referred
to \textbf{black-box} models. Despite their shortcomings, black-box
models have grown in popularity in recent years and have at times
created undesirable societal outcomes
\protect\hyperlink{ref-o2016weapons}{{[}2{]}}. The scientific community
has tackled this issue from two different angles: while some have
appealed for a strict focus on inherently interpretable models
\protect\hyperlink{ref-rudin2019stop}{{[}3{]}}, others have investigated
different ways to explain the behaviour of black-box models. These two
sub-domains can be broadly referred to as \textbf{interpretable AI} and
\textbf{explainable AI} (XAI), respectively.

Among the approaches to XAI that have recently grown in popularity are
\textbf{Counterfactual Explanations} (CE). They explain how inputs into
a model need to change for it to produce different outputs.
Counterfactual Explanations that involve realistic and actionable
changes can be used for the purpose of \textbf{Algorithmic Recourse}
(AR) to help individuals who face adverse outcomes. An example relevant
to the Social Sciences is consumer credit: in this context, AR can be
used to guide individuals in improving their creditworthiness, should
they have previously been denied access to credit based on an automated
decision-making system. A meaningful recourse recommendation for a
denied applicant could be: \emph{``If your net savings rate had been
10\% of your monthly income instead of the actual 8\%, your application
would have been successful. See if you can temporarily cut down on
consumption.''} In the remainder of this paper we will use both
terminologies - recourse and counterfactual - interchangeably to refer
to situations where counterfactuals are generated with the intent to
provide individual recourse.

Existing work in this field has largely worked in a static setting:
various approaches have been proposed to generate counterfactuals for a
given individual that is subject to some pre-trained model. More recent
work has compared different approaches within this static setting
\protect\hyperlink{ref-pawelczyk2021carla}{{[}4{]}}. In this work, we go
one step further and ask ourselves: what happens if recourse is provided
and implemented repeatedly? What types of dynamics are introduced and
how do different counterfactual generators compare in this context?

Research on Algorithmic Recourse has also so far typically addressed the
issue from the perspective of one single individual and has indeed been
referred to as \textbf{individual recourse} in some places. Arguably
though, most real-world applications that warrant Algorithmic Recourse
involve potentially large groups of individuals typically competing for
scarce resources. Our work demonstrates that in such scenarios, choices
made by or for one single individual are likely to affect the broader
collective of individuals in ways that current approaches to AR fail to
account for. More specifically, we argue that a strict focus on
minimizing the private costs faced by individuals may be too narrow an
objective.

Figure~\ref{fig-poc} illustrates this idea for a binary problem
involving a probabilistic classifier and the counterfactual generator
proposed by \protect\hyperlink{ref-wachter2017counterfactual}{{[}5{]}}:
the implementation of AR for a subset of individuals leads to a domain
shift (b), which in turn triggers a model shift (c). As this game of
implementing AR and updating the classifier is repeated, the decision
boundary moves away from training samples that were originally in the
target class (d). We refer to these types of dynamics as
\textbf{endogenous} because they are induced by the implementation of
recourse itself. The term \textbf{macrodynamics} is borrowed from the
economics literature and used to describe processes involving whole
groups or societies.

\begin{figure}

{\centering \includegraphics[width=0.45\textwidth,height=\textheight]{www/poc.png}

}

\caption{\label{fig-poc}Dynamics in Algorithmic Recourse: we have a
simple Bayesian model trained for binary classification (a); the
implementation of AR for a random subset of individuals leads to a
domain shift (b); as the classifier is retrained we observe a model
shift (c); as this process is repeated, the decision boundary moves away
from the target class (d).}

\end{figure}

We think that these types of endogenous dynamics may be problematic and
warrant our attention. Firstly, model shifts may inadvertently change
classification outcomes for individuals who never received and
implemented recourse. Secondly and relatedly, we observe in
Figure~\ref{fig-poc} that as the decision boundary moves in the
direction of the non-target class, counterfactual paths become shorter:
in the consumer credit example, individuals that previously would have
been denied credit based on their input features are suddenly considered
as creditworthy. Average default risk across all borrowers can therefore
be expected to increase. Conversely, lenders that anticipate such
dynamics may choose to refrain from offering recourse (and hence credit)
to more than just a tiny share of individuals. In that latter and
perhaps more likely scenario, the probability of being offered recourse
decreases with every individual that implements recourse: in other
words, the actions of first-movers exert a negative externality on
future would-be borrowers.

To the best of our knowledge this is the first work investigating
endogenous macrodynamics in AR. Our contributions to the state of
knowledge are as follows: firstly, we posit a compelling argument that
calls for a novel perspective on Algorithmic Recourse extending our
focus from single individuals to groups. Secondly, we introduce an
experimental framework extending previous work by
\protect\hyperlink{ref-altmeyer2022CounterfactualExplanations}{{[}6{]}},
which enables us to study macrodynamics of Algorithmic Recourse through
simulations that can be fully parallelized. Thirdly, we use this
framework to provide a first in-depth analysis of endogenous recourse
dynamics induced by various popular counterfactual generators including
\protect\hyperlink{ref-wachter2017counterfactual}{{[}5{]}},
\protect\hyperlink{ref-schut2021generating}{{[}7{]}},
\protect\hyperlink{ref-joshi2019towards}{{[}8{]}},
\protect\hyperlink{ref-mothilal2020explaining}{{[}9{]}} and
\protect\hyperlink{ref-antoran2020getting}{{[}10{]}}. To this end we
propose a number of novel evaluation metrics that can be used to
quantify and benchmark the macrodynamics introduced by the different
generators. Finally, we also discuss what drives endogenous dynamics and
propose strategies to mitigate them.

The remainder of the paper is structured as follows:
Section~\ref{sec-related} places our work in the broader context of
related literature. Section~\ref{sec-method} posits a generalized
framework for gradient-based Algorithmic Recourse and introduces the
notion of hidden external costs. Section~\ref{sec-method-2} sets out our
experimental framework for modeling endogenous macrodynamics in AR.
Section~\ref{sec-empirical} presents our experimental results for
synthetic and real-world datasets. We also present evidence for the
effectiveness of various proposed mitigation strategies in that section.
Our findings are then discussed in the broader context of the literature
in Section~\ref{sec-discussion}. We also point to some of the
limitations of our work as well as avenues for future research in
Section~\ref{sec-limit}. Finally, Section~\ref{sec-conclusion}
concludes.

\hypertarget{sec-related}{%
\section{Background}\label{sec-related}}

In this Section we provide a review of the relevant literature. First,
Section~\ref{sec-related-recourse} discusses the existing research
within the domain of Counterfactual Explanations and Algorithmic
Recourse. Then, Section~\ref{sec-related-shifts} presents some of the
previous work on the measurement of dataset and model shifts.

\hypertarget{sec-related-recourse}{%
\subsection{Algorithmic Recourse}\label{sec-related-recourse}}

A framework for Counterfactual Explanations was first proposed in 2017
by \protect\hyperlink{ref-wachter2017counterfactual}{{[}5{]}} and has
served as the baseline for most methodologies that have been proposed
since then. Let \(M: \mathcal{X} \mapsto \mathcal{Y}\) denote some
pre-trained model that maps from inputs \(X \in \mathcal{X}\) to outputs
\(Y \in \mathcal{Y}\). Then we are interested in minimizing the
cost\footnote{Equivalently, others have referred to this quantity as
  \emph{complexity} or simply \emph{distance}} \(H=h(x\prime)\) incurred
by individual \(x\) when moving to a counterfactual state \(x\prime\)
such that the predicted outcome \(M(x\prime)\) corresponds to some
target outcome \(y^*\):

\begin{equation}\protect\hypertarget{eq-obj}{}{
\min_{x\prime \in \mathcal{x}} h(x\prime) \ \ \ \mbox{s. t.} \ \ \ M(x\prime) = t
}\label{eq-obj}\end{equation}

For implementation purposes, Equation~\ref{eq-obj} is typically
approximated through regularization:

\begin{equation}\protect\hypertarget{eq-solution}{}{
x\prime = \arg \min_{x\prime}  \ell(M(x\prime),y^*) + \lambda h(x\prime)
}\label{eq-solution}\end{equation}

In the baseline work
\protect\hyperlink{ref-wachter2017counterfactual}{{[}5{]}}, the cost
function \(h: \mathcal{X} \mapsto \mathbb{R}\) is proxied by some
distance metric based on the simple intuition that perturbations of
\(x\) are costly to the individual. For models that are differentiable
and produce smooth predictions, Equation~\ref{eq-solution} can be solved
through gradient descent. This summarizes the approach followed in
\protect\hyperlink{ref-wachter2017counterfactual}{{[}5{]}} which we
shall refer to simply as \textbf{Wachter} - the name of the first author
- in the remainder of this paper.

Many approaches for the generation of Algorithmic Recourse have been
described in the literature since 2017. An October 2020 survey by Karimi
et al.~layed out 60 algorithms that have been proposed since 2014
\protect\hyperlink{ref-karimi2020survey}{{[}11{]}}. Another survey
published around the same time by Verma et al.~described 29 algorithms
\protect\hyperlink{ref-verma2020counterfactual}{{[}12{]}}. Different
approaches vary primarily in terms of the objective functions they
impose, how they optimize said objective (from brute force through
gradient-based approaches to graph traversal algorithms), and how the
ensure that certain requirements for CE are met. Regarding the latter,
the literature has produced an extensive list of desiderata each
addressing different needs. To name but a few, we are interested in
generating counterfactuals that close
\protect\hyperlink{ref-wachter2017counterfactual}{{[}5{]}}, actionable
\protect\hyperlink{ref-ustun2019actionable}{{[}13{]}}, realistic
\protect\hyperlink{ref-schut2021generating}{{[}7{]}}, sparse, diverse
\protect\hyperlink{ref-mothilal2020explaining}{{[}9{]}} and if possible
causally founded
\protect\hyperlink{ref-karimi2021algorithmic}{{[}14{]}}.

Efforts so far have largely been directed at improving the quality of
Counterfactual Explanations within a static context: given some
pre-trained classifier \(M: \mathcal{X} \mapsto \mathcal{Y}\), we are
interested in generating one or multiple meaningful Counterfactual
Explanations for some individual characterized by \(x\). The ability of
Counterfactual Explanations to handle dynamics like data and model
shifts remains a largely unexplored research challenge at this point
\protect\hyperlink{ref-verma2020counterfactual}{{[}12{]}}. We have been
able to identify only one recent work that considers the implications of
\textbf{exogenous} domain and model shifts in the context of AR
\protect\hyperlink{ref-upadhyay2021towards}{{[}15{]}}. Exogenous shifts
are strictly of external origin. For example, they might stem from data
correction, temporal shifts or geospatial changes
\protect\hyperlink{ref-upadhyay2021towards}{{[}15{]}}. The authors of
\protect\hyperlink{ref-upadhyay2021towards}{{[}15{]}} propose ROAR - a
framework for Algorithmic Recourse that evidently improves robustness to
such exogenous shifts.

As mentioned earlier, research has so far also generally focused on
generating counterfactuals for single individuals or instances. We have
been able to identify only one existing work that investigates black-box
model behavior towards a group of individuals
\protect\hyperlink{ref-carrizosa2021generating}{{[}16{]}}. The authors
propose an optimization framework that generates collective
counterfactuals. We provide a motivation for doing so from the
perspective of endogenous macrodynamics of Algorithmic Recourse.

\hypertarget{sec-related-shifts}{%
\subsection{Domain and Model Shifts}\label{sec-related-shifts}}

Much attention has been paid to the detection of dataset shifts --
situations where the distribution of data changes over time. Rabanser et
al.~suggest a framework to detect data drift from a minimal number of
samples through the application of two-sample tests
\protect\hyperlink{ref-rabanser2019failing}{{[}17{]}}. This task is a
generalization of the anomaly detection problem for large datasets,
which aims to answer the question if two sets of samples could have been
generated from the same probability distribution. Numerous approaches to
anomaly detection have been summarized
\protect\hyperlink{ref-chandola2009anomaly}{{[}18{]}}. Another
well-established research topic is that of concept drift -- situations
where external variables influence the patterns between the input and
the output of a model
\protect\hyperlink{ref-widmer1996learning}{{[}19{]}}. For instance, Gama
et al.~offer a review of the adaptive learning techniques which can
handle concept drift \protect\hyperlink{ref-gama2014survey}{{[}20{]}}.
Less previous work is available on the related topic of model drift -
changes in model performance over time. Nelson et al.~review how
resistant different machine learning models are to the model drift
\protect\hyperlink{ref-nelson2015evaluating}{{[}21{]}}. Ackerman et
al.~offer a method to detect changes in model performance when ground
truth is not available
\protect\hyperlink{ref-ackerman2021machine}{{[}22{]}}.

In the context of Algorithmic Recourse, domain and model shifts were
first brought up by the authors behind ROAR
\protect\hyperlink{ref-upadhyay2021towards}{{[}15{]}}. In their work
they refer to model shifts as simply any perturbation \(\Delta\) to the
parameters of the model in question: \(M\). While this also sets the
baseline for our analysis here, it is worth noting that in
\protect\hyperlink{ref-upadhyay2021towards}{{[}15{]}} these
perturbations are mechanically introduced. In contrast we are interested
in quantifying model shifts that arise endogenously as part of a dynamic
recourse process. In addition to quantifying the magnitude of shifts
\(\Delta\), we aim to also analyse the characteristics of changes to the
model, such as the position of the decision boundary and the overall
decisiveness of the model. We have not been able to identify previous
work on this topic.

\hypertarget{sec-related-benchmark}{%
\subsection{Benchmarking Counterfactual
Generators}\label{sec-related-benchmark}}

Despite the large and growing number of approaches to counterfactual
search, there have been surprisingly few benchmark studies that compare
different methodologies. This may be partially due to limited software
availability in this space. Recent work has started to address this gap:
firstly, \protect\hyperlink{ref-de2021framework}{{[}23{]}} run a large
benchmarking study using different algorithmic aproaches and numerous
tabular datasets; secondly,
\protect\hyperlink{ref-pawelczyk2021carla}{{[}4{]}} introduce a Python
framework - CARLA - that can be used to apply and benchmark different
methodologies; finally, \texttt{CounterfactualExplanations.jl}
\protect\hyperlink{ref-altmeyer2022CounterfactualExplanations}{{[}6{]}}
provides an extensible, fast and language-agnostic implementation in
Julia. Since the experiments presented here involve extensive
simulations, we have relied on and extended the Julia implementation due
to the associated performance benefits. In particular, we have built a
framework on top of \texttt{CounterfactualExplanations.jl} that extends
the functionality from static benchmarks to simulation experiments:
\texttt{AlgorithmicRecourseDynamics.jl}\footnote{The package is
  available from \ldots{}}. The core concepts implemented in that
package reflect what is presented in section Section~\ref{sec-method-2}
of this paper.

\hypertarget{sec-method}{%
\section{Gradient-Based Recourse Revisited}\label{sec-method}}

In this section we first set out a generalized framework for
gradient-based counterfactual search in Section~\ref{sec-method-general}
that encapsulates the various individual recourse methods we have chosen
to use in our experiments. We then introduce the notion of a hidden
external cost in algorithmic recourse and extend the existing framework
to explicitly address this cost in the counterfactual search objective.

\hypertarget{sec-method-general}{%
\subsection{From individual recourse
\ldots{}}\label{sec-method-general}}

We have chosen to focus on gradient-based counterfactual search for two
reasons: firstly, they can be seen as direct descendants of our baseline
method - Wachter; secondly, gradient-based search is particularly
well-suited for differentiable black-box models like deep neural
networks, which we focus on in this work. In particular, we include
include the following generators in our simulation experiments below:
\textbf{REVISE} \protect\hyperlink{ref-joshi2019towards}{{[}8{]}},
\textbf{CLUE} \protect\hyperlink{ref-antoran2020getting}{{[}10{]}},
\textbf{DiCE} \protect\hyperlink{ref-mothilal2020explaining}{{[}9{]}}
and a greedy approach that relies on probabilistic models
\protect\hyperlink{ref-schut2021generating}{{[}7{]}}. Our motivation for
including these different generators in our analysis, is that they all
offer slightly different approaches to generate meaningful
counterfactuals for differentiable black-box models. We hypothesize that
generating more \textbf{meaningful} counterfactuals should mitigate the
endogenous dynamics illustrated in Figure~\ref{fig-poc} in
Section~\ref{sec-intro}. This intuition stems from the underlying idea
that more meaningful counterfactuals are generated by the same or at
least a very similar data generating process as the training data. All
else equal, counterfactuals that fulfill this basic requirement should
be less prone to trigger domain and model shifts.

As we will see next, all of them can be described by the following
generalized form of Equation~\ref{eq-solution}:

\begin{equation}\protect\hypertarget{eq-general}{}{
\begin{aligned}
\mathbf{s}^\prime &= \arg \min_{\mathbf{s}^\prime \in \mathcal{S}} \left\{ \sum_{k=1}^{K} {\ell(M(f(s_k^\prime)),y^*)}+ \lambda {h(f(s_k^\prime)) }  \right\}
\end{aligned}
}\label{eq-general}\end{equation}

Here \(\mathbf{s}^\prime=\left\{s_k^\prime\right\}_K\) is the stacked
\(K\)-dimensional array of counterfactual states and
\(f: \mathcal{S} \mapsto \mathcal{X}\) maps from the counterfactual
state space to the feature space. In Wachter, the state space is the
feature space: \(f\) is just the identity function and the number of
counterfactuals \(K\) is equal to one. Both REVISE and CLUE search
counterfactuals in some latent embedding \(S \subset \mathcal{S}\)
instead of the feature space directly. The latent embedding is learned
by a separate generative model that is tasked with learning the data
generating process (DGP) of \(X\). In this case \(f\) in
Equation~\ref{eq-general} corresponds to the decoder part of the
generative model, in other words the function that maps back from the
latent embedding to the feature space. Provided the generative model is
well-specified, traversing the latent embedding typically results in
realistic and plausible counterfactuals, because they are implicitly
generated by the (learned) DGP
\protect\hyperlink{ref-joshi2019towards}{{[}8{]}}.

CLUE distinguishes itself from REVISE and other counterfactual
generators in that it aims to minimize the predictive uncertainty of the
model in question, \(M\). To quantify predictive uncertainty the authors
rely on entropy estimates for probabilistic models. The approach
proposed by \protect\hyperlink{ref-schut2021generating}{{[}7{]}}, which
we shall refer to as \textbf{Greedy}, also works with the subclass of
models \(\tilde{\mathcal{M}}\subset\mathcal{M}\) that can produce
predictive uncertainty estimates. The authors show that in this setting
the cost function \(h(\cdot)\) in Equation~\ref{eq-general} is redundant
and meaningful counterfactuals can be generated in a fast and efficient
manner through a modified Jacobian-based Saliency Map Attack (JSMA).
Schut et al. \protect\hyperlink{ref-schut2021generating}{{[}7{]}} also
show that by maximizing the predicted probability of \(x^\prime\) being
assigned to target class \(y^*\) we also implicitly minimize predictive
entropy - as in CLUE. In that sense, CLUE can be seen as equivalent to
REVISE in the Bayesian context and we shall therefore refer to both
approaches collectively as \textbf{Latent Space} generators\footnote{In
  fact, there are a number of other recently proposed approaches to
  counterfactual search that also broadly fall in this same category.
  They largely differ with respect to the chosen generative model: for
  example, the generator proposed by
  \protect\hyperlink{ref-dombrowski2021diffeomorphic}{{[}24{]}} relies
  on normalizing flows.}.

Finally, DiCE \protect\hyperlink{ref-mothilal2020explaining}{{[}9{]}}
distinguishes itself from all other generators considered here in that
it aims to generate a diverse set of \(K>1\) counterfactuals. Wachter et
al. \protect\hyperlink{ref-wachter2017counterfactual}{{[}5{]}} show that
diverse outcomes can in principal be achieved simply rerunning
counterfactual search multiple times using stochastic gradient descent
(or by randomly initializing the counterfactual). In
\protect\hyperlink{ref-mothilal2020explaining}{{[}9{]}} diversity is
explicitly proxied via Determinantal Point Processes (DDP): the authors
simply introduce DDP as a component of the cost function
\(h(\mathbf{s}^\prime)\) and therebt produce counterfactuals
\(s_1, ... , s_K\) that look as different from each other as possible.
The implementation of DiCE in the our library of choice -
\texttt{CounterfactualExplanations.jl} - uses that exact approach. It is
worth noting that for \(k=1\), DiCE reduces to Wachter since the DDP is
constant and therefore does not affect the objective function
Equation~\ref{eq-general}.

\hypertarget{towards-collective-recourse}{%
\subsection{\ldots{} towards collective
recourse}\label{towards-collective-recourse}}

All of the different approaches introduced above tackle the problem of
Algorithmic Recourse from the perspective of one single
individual\footnote{DiCE recognizes that different individuals may have
  different objective functions, but it does not address the
  interdependencies between different individuals.}. To explicitly
address the issue that individual recourse may affect the outcome and
prospect of other individuals, we propose to extend
Equation~\ref{eq-general} as follows:

\begin{equation}\protect\hypertarget{eq-collective}{}{
\begin{aligned}
\mathbf{s}^\prime &= \arg \min_{\mathbf{s}^\prime \in \mathcal{S}}  \sum_{k=1}^{K} {\ell(M(f(s_k^\prime)),y^*)} \\ &+ \lambda_1 {h(f(s_k^\prime)) } + \lambda_2 {g(f(s_k^\prime))}  
\end{aligned}
}\label{eq-collective}\end{equation}

Here \(h(f(s_k^\prime))\) denotes the proxy for private costs faced by
the individual as before and \(\lambda_1\) governs to what extent that
private cost ought to be penalized. The newly introduced term
\(g(f(s_k^\prime))\) is meant to capture and address external costs
incurred by the collective of individuals in response to changes in
\(\mathbf{s}^\prime\). The underlying concept of private and external
costs is borrowed from Economics and well-established in that field:
when the decisions or actions by some individual market participant
generate external costs, then the market is said to suffer from negative
externalities and considered inefficient
\protect\hyperlink{ref-pindyck2014microeconomics}{{[}25{]}}. We think
that this concept describes the endogenous dynamics of algorithmic
recourse oberserved here very well. As with individual recourse, the
exact choice of \(g(\cdot)\) is not obvious, nor do we intend to provide
a definite answer in this work, if such even exists. That being said, we
do propose a few potential mitigation strategies in
Section~\ref{sec-empirical-2-mitigate}.

\hypertarget{sec-method-2}{%
\section{Modeling Endogenous Macrodynamics in Algorithmic
Recourse}\label{sec-method-2}}

In the following we describe the framework we propose for modeling and
analysing endogenous macrodynamics in Algorithmic Recourse. We first
describe the basic simulations that were generated to produce the
findings in this work and also constitute the core of
\texttt{AlgorithmicRecourseDynamics.jl} - the Julia package we
introduced earlier. The remainder of this section then introduces
various evaluation metrics that can be used to benchmark different
counterfactual generators with respect to how they perform in the
dynamic setting.

\hypertarget{sec-method-2-experiment}{%
\subsection{Simulations}\label{sec-method-2-experiment}}

The dynamics illustrated in Figure~\ref{fig-poc} in
Section~\ref{sec-intro} were generated through a simple experiment that
aims to simulate the process of Algorithmic Recourse in practice. We
begin in the static setting at time time \(t=0\): given some pre-trained
classifier \(M\) we generate recourse for a random batch of \(B\)
individuals in the non-target class. Note that we focus our attention on
classification problems, since classification poses the most common
practical use-case for Algorithmic Recourse. In order to simulate the
dynamic process, we suppose that the model \(M\) is retrained following
the actual implementation of recourse in time \(t=0\). Following the
update to the model, we assume that at time \(t=1\) recourse is
generated for yet another random subset of individuals in the non-target
class. This process is repeated for a number of time periods \(T\). To
get a clean read on endogenous dynamics we keep the total population of
samples closed: we allow existing samples to move from factual to
counterfactual states, but do not allow any entirely new samples to
enter the population. The experimental setup is summarized in Algorithm
\ref{algo-experiment}

\begin{algorithm}
\caption{Simulation Experiment}\label{algo-experiment}
\begin{algorithmic}[1]
\Procedure{Experiment}{$M,D,G$}
\State $t\gets 0$
\While{$t<T$}
\State $D_B \subset D$
\State $D_B\gets G(D_B)$ \Comment{Generate counterfactuals.}
\State $M\gets M(D)$ \Comment{Retrain model.}
\EndWhile
\State \textbf{return} $M,D$
\EndProcedure
\end{algorithmic}
\end{algorithm}

A noteworthy practical consideration is the choice of \(T\) and \(B\).
The higher these values, the more factual instances undergo recourse
throughout the entire experiment. Of course, this is likely to lead to
more pronounced domain and model shifts by time \(T\). At the same time,
it is generally improbable that a very large part of the population
would request an explanation of the algorithm's decisions. In our
experiments, we choose the values such that \(T \cdot B\) corresponds to
the application of recourse on \(25-50\%\) of the negative instances
from the initial dataset. As we collect data at each time \(t\), we can
also verify the impact of recourse when it is implemented for a smaller
number of individuals.

Algorithm \ref{algo-experiment} summarizes the proposed simulation
experiment for a given dataset \(D\), model \(M\) and generator \(G\),
but naturally we are interested in comparing simulation outcomes for
different sources of data, models and generators. The framework we have
built facilitates this, making use of multi-threading in order to speed
up computations. Holding the initial model and dataset constant the
experiments are run for all generators, since our primary concern is to
benchmark different recourse methods. To ensure that each generator is
faced with the exact same initial conditions in each round \(t\), the
candidate batch of individuals from the non-target class is randomly
drawn from the intersection of all individuals in the non-target class
across all experiments
\(\left\{\textsc{Experiment}(M,D,G)\right\}_{j=1}^J\) where \(J\) is the
total number of generators.

\hypertarget{sec-method-2-metrics}{%
\subsection{Evaluation Metrics}\label{sec-method-2-metrics}}

We formulate two desiderata for the set of metrics used to measure
domain and model shifts induced by recourse. First, the metrics should
be applicable regardless of the dataset or classification technique so
that they allow for the meaningful comparison of the generators in
various scenarios. As the knowledge of the underlying probability
distribution is rarely available, the metrics should be empirical and
non-parametric. This further ensures that we can also measure large
datasets by sampling from the available data. Moreover, while our study
was conducted in a two-class classification setting, our choice of
metrics should remain applicable in the future research on multi- class
recourse problems. Second, the set of metrics should allow to capture
various aspects of the previously mentioned magnitude, path, and tempo
of changes while remaining as small as possible.

\hypertarget{domain-shifts}{%
\subsubsection{Domain Shifts}\label{domain-shifts}}

To quantify the magnitude of domain shifts we rely on an unbiased
estimate of the squared population \textbf{Maximum Mean Discrepancy
(MMD)} given as:

\begin{equation}\protect\hypertarget{eq-mmd}{}{
\begin{aligned}
MMD^2_u[F,{X}^\prime,\tilde{X}^\prime] &= \frac{1}{m(m-1)}\sum_{i=1}^m\sum_{j\neq i}^m k(x_i,x_j) \\ &+ \frac{1}{n(n-1)}\sum_{i=1}^n\sum_{j\neq i}^n k(\tilde{x}_i,\tilde{x}_j) \\ &- \frac{2}{mn}\sum_{i=1}^m\sum_{j=1}^n k(x_i,\tilde{x}_j)
\end{aligned}
}\label{eq-mmd}\end{equation}

where \(\mathcal{F}\) is a unit ball in a Reproducing Kernel Hilbert
Space H {[}27{]}, and \(X\), \(\tilde{X}\) represent independent and
identically distributed samples drawn from probability distributions
\(p\) and \(q\) respectively {[}28{]}. MMD is a measure of the distance
between the kernel mean embeddings of \(p\) and \(q\) in RKHS
\(\mathcal{H}\). An important consideration is the choice of the kernel
function \(k(\cdot,\cdot)\). In our implementation we make use of the
radial basis function (RBF) kernel with a constant length-scale
parameter of \(0.5\). As RBF captures all moments of distributions \(p\)
and \(q\), we have that \(MMD_u^2[F,X,\tilde{X}]=0\) if and only if
\(X=\tilde{X}\).

The evaluation metric in Equation~\ref{eq-mmd} is computed after every
round \(t=1,...,T\) of the experiment. To assess the statistical
significance of the observed shifts under the null hypothesis that
samples \(X\) and \(\tilde{X}\) were drawn from the same probability
distribution we follow
\protect\hyperlink{ref-arcones1992bootstrap}{{[}26{]}}. To that end, we
combine the two samples and generate a large number of permutations of
\(X + \tilde{X}\). Then, we split the permuted data into two new samples
\(X^\prime\) and \(\tilde{X}^\prime\) having the same size as the
original samples. Then under the null hypothesis we should have that
\(MMD_u^2[F,X^\prime,\tilde{X}^\prime]\) be approximately equal to
\(MMD_u^2[F,X,\tilde{X}]\). The corresponding \(p\)-value can then be
calculated by counting how these two quantities are not equal.

We calculate the MMD for both classes individually based on the ground
truth labels, i.e.~the labels that samples were assigned in time
\(t=0\). Throughout our experiments we generally do not expect the
distribution of the negative class to change over time -- application of
recourse reduces the size of this class, but since individuals are
sampled uniformly the distribution should remain unaffected. Conversely,
unless a recourse generator can perfectly replicate the original
probability distribution, we expect the MMD of the positive class to
increase. Thus, when discussing MMD, we generally mean the shift in the
distribution of the positive class.

Finally, \textbf{feature mean and feature standard deviation} are also
calculated to verify how the implementation of recourse impacts every
attribute in the dataset. Although MMD already captures information
about the expected value and variance, we may also be interested in a
more granular look at individual features.

\hypertarget{model-shifts}{%
\subsubsection{Model Shifts}\label{model-shifts}}

As our baseline for quantifying model shifts we measure perturbations to
the model parameters at each point in time \(t\) following
\protect\hyperlink{ref-upadhyay2021towards}{{[}15{]}}. We define
\(\Delta=||\theta_{t+1}-\theta_{t}||^2\), that is the euclidean distance
between the vectors of parameters before and after retraining the model
\(M\). We shall refer to this baseline metric simply as
\textbf{Perturbations}.

We extend the metric in Equation~\ref{eq-mmd} for the purpose of
quantifying model shifts. Specifically, we introduce \textbf{Predicted
Probability MMD (PP MMD)}: instead of applying Equation~\ref{eq-mmd} to
features directly, we apply it to the predicted probabilities assigned
to a set of samples by the model \(M\). If the model shifts, the
probabilities assigned to each sample will change; again, this metric
will equal 0 only if the two classifiers are the same. It is worth
noting that while we apply the technique to samples drawn uniformly from
the dataset, it can also be employed on arbitrary points in the entire
feature space (or a subspace). The latter approach is theoretically more
robust. Unfortunately, in practice this approachs suffers from the curse
of dimensionality, since it becomes increasingly difficult to select
enough points to overcome noise as the dimension \(D\) grows.

As an alternative to PP MMD we use a pseudo-distance for the
\textbf{Disagreement Coefficient} (Disagreement). This metric was
introduced in \protect\hyperlink{ref-hanneke2007bound}{{[}27{]}} and
estimates \(p(M(x) \neq M^\prime(x))\), that is the probability that two
classifiers do not agree on the predicted outcome for a randomly chosen
sample. Thus, it is not relevant whether the classification is correct
according to the ground truth, but only whether the sample lies on the
same side of the two respective decision boundaries. In our context,
this metric quantifies the overlap between the initial model (trained
before the application of recourse) and the updated model. A
Disagreement Coefficient unequal to zero is indicative of a model shift.
The opposite is not true: even if the Disagreement Coefficient is equal
to zero a model shift may still have occured. This is one reason for why
PP MMD is our our preferred metric.

Finally, we introduce \textbf{Decisiveness} as a metric that quantifies
the likelihood that a model assigns a high probability to its
classification of any given sample. We define the metric simply as
\({1\over{N}}\sum_{i=0}^N(\sigma(M(x)) − 0.5)^2\) where \(M(x)\) are
predicted logits from a binary classifier and \(\sigma\) denotes the
sigmoid function. This metric provides an unbiased estimate of the
binary classifier's tendency to produce high-confidence predictions in
either one of the two classes. Although the exact values for this metric
are not important for our study, they can be used to detect model
shifts. If decisiveness changes over time, then this is indicative of
the decision boundary moves towards either one of the two classes.

\hypertarget{sec-empirical}{%
\section{Experiment Setup}\label{sec-empirical}}

\hypertarget{sec-empirical-data}{%
\subsection{Data}\label{sec-empirical-data}}

We have chosen to work with both synthetic and real-world datasets.
Using synthetic data allows us to impose distributional properties that
may affect the resulting recourse dynamics. Following
\protect\hyperlink{ref-upadhyay2021towards}{{[}15{]}}, we generate
synthetic data in \(\mathbb{R}^2\) to also allow for a visual
interpretation of the results. Real-world data is used in order to
assess if endogenous dynamics also occur in higher-dimensional settings.

\hypertarget{synthetic-data}{%
\subsubsection{Synthetic data}\label{synthetic-data}}

We use four synthetic binary classification datasets consisting of 1000
samples each. The datasets are presented in
Figure~\ref{fig-synthetic-data} (see also Appendix A for a formal
description). Samples from the negative class are marked in blue while
samples of the positive class are marked in orange.

\begin{figure}

{\centering \includegraphics[width=8cm,height=2cm]{www/synthetic_data.png}

}

\caption{\label{fig-synthetic-data}Synthetic classification datasets
used in our experiments.}

\end{figure}

Ex-ante we expect to see that by construction Wachter will create a new
cluster of counterfactual instances in the proximity of the initial
decision boundary. Thus, the choice of a black-box model may have an
impact on the paths of the recourse. For generators that use latent
space search (REVISE \protect\hyperlink{ref-joshi2019towards}{{[}8{]}},
CLUE \protect\hyperlink{ref-antoran2020getting}{{[}10{]}}) or rely on
(and have access to) probabilistic models (CLUE
\protect\hyperlink{ref-antoran2020getting}{{[}10{]}}, Greedy
\protect\hyperlink{ref-schut2021generating}{{[}7{]}}) we expect that
counterfactuals will end up in regions of the target domain that are
densely populated by training samples. Of course, this is expectation
hinges on how effective said probabilistic models are at capturing
predictive uncertainty. Finally, we expect to see the counterfactuals
generated by DiCE to be uniformly spread around the feature space inside
the target class\footnote{As we mentioned earlier, the diversity
  constraint used by DiCE is only effective for when at least two
  counterfactuals are being generated. We have therefore decided to
  always generate 5 counterfactuals for each generator and randomly pick
  one of them.}. In summary, we expect that the endogenous shifts
induced by Wachter outsize those induced by all other generators, since
Wachter is the only approach that is not concered with generating what
we have defined as meaningful counterfactuals.

\hypertarget{real-world-data}{%
\subsubsection{Real-world data}\label{real-world-data}}

We use three different real-world datasets from the Finance and
Economics domain, all of which are tabular and can be used for binary
classification. Firstly, we use the \textbf{Give Me Some Credit} dataset
which was open-sourced on Kaggle for the task to predict whether a
borrower is likely to experience financial difficulties in the next two
years \protect\hyperlink{ref-gmsc_data}{{[}28{]}}. Originally consisting
of 250,000 instances with 11 numerical attributes. Secondly, we use the
\textbf{UCI defaultCredit} dataset
\protect\hyperlink{ref-yeh2009comparisons}{{[}29{]}}, a benchmark
dataset that can be used to train binary classifiers to predict the
binary outcome variable, whether credit card clients default on their
payment. In its raw form it consists of 23 explanatory variables - 4
categorical features relating to demographic attributes\footnote{These
  have been ommitted from the analysis. See Section~\ref{sec-limit-data}
  for details.} and 19 continuous features largely relating to
individuals' payment histories and amount of credit outstanding. Both of
these datasets have been used in the literature on Algorithmic Recourse
before (see for example
\protect\hyperlink{ref-pawelczyk2021carla}{{[}4{]}},
\protect\hyperlink{ref-joshi2019towards}{{[}8{]}} and
\protect\hyperlink{ref-ustun2019actionable}{{[}13{]}}), presumably
because they constitute real-world classification tasks involving
individuals that compete for access to credit.

As a third dataset we include the \textbf{California Housing} dataset
derived from the 1990 U.S. census and sourced through scikit-learn
\protect\hyperlink{ref-pace1997sparse}{{[}31{]}}. It consists of 8
continuous features that can be used to predict the median house price
for California districts. The continuous outcome variable is binarized
as \(\tilde{y}=\mathbb{I}_{y>\text{median}(Y)}\) indicating whether or
not the median house price of a given district is above or below the
median of all districts. While we have not seen this dataset used in the
previous literature on AR, others have used the Boston Housing dataset
in a similar fashion (see for example
\protect\hyperlink{ref-schut2021generating}{{[}7{]}}). While we
initially also conducted experiments on that dataset, we eventually
discarded this dataset, since it has been found to suffer from an
ethical problem \protect\hyperlink{ref-carlisle2019racist}{{[}32{]}}.

Since the simulations involve generating counterfactuals for a
significant proportion of the entire sample of individuals, we have
randomly undersampled each dataset to yield balanced subsamples
consisting of 10,000 individuals each. We have also standardized all
explanatory features since our chosen classifiers are sensetive to
scale.

\hypertarget{sec-empirical-classifiers}{%
\subsection{Classifiers and Generative
Models}\label{sec-empirical-classifiers}}

For each dataset and generator we look at three different types of
classifiers all of them built and trained using \texttt{Flux.jl}
\protect\hyperlink{ref-innes2018fashionable}{{[}33{]}}: firstly, a
simple linear classifier - \textbf{Logistic Regression} - implemented as
single linear layer with sigmoid activation; secondly, a multilayer
perceptron (\textbf{MLP}); and finally, a \textbf{Deep Ensemble}
composed of five MLPs following
\protect\hyperlink{ref-lakshminarayanan2016simple}{{[}34{]}} that serves
as our only probabilistic classifier. We have chosen to work with deep
ensembles both for their simplicity and effectiveness at modelling
predictive uncertainty. They are also the model of choice in
\protect\hyperlink{ref-schut2021generating}{{[}7{]}}. The actual neural
network architectures are kept simple (Table~\ref{tbl-mlp}), since we
are only marginally concerned with achieving good initial classifier
performance. For the real-world datasets we using mini-batch training
and dropout regularization.

The Latent Space generators rely on separate generative models.
Following the authors of both REVISE and CLUE we use Variational
Autoencoders (\textbf{VAE}) for this purpose. As with the classifiers,
we deliberately choose to work with fairly simple architectures
(Table~\ref{tbl-vae}). More expressive generative models generally also
lead to more meaningful counterfactuals produced by Latent Space
generators. But in our view this should simply be considered as a
vulnerability of counterfactual generators that rely on surrogate models
to learn what realistic representations of the underlying data.

All classifiers and generative models are retrained for 10 epochs in
each round \(t\) of the experiment. Rather than retraining models from
scratch, we initialize all parameters at their previous levels (\(t-1\))
and compute backpropagate for 10 epochs using the new training data as
inputs into the existing model.

\begin{table}

\caption{\label{tbl-panel}Model
Architectures}\begin{minipage}[t]{\linewidth}
\subcaption{\label{tbl-mlp}MLP}

{\centering

\begin{tabular}[t]{lllll}
\toprule
 & Hidden Dim. & Hidden Layers & Batch & Dropout\\
\midrule
Synthetic & 32 & 1 & - & -\\
Real-World & 32 & 2 & 50 & 0.25\\
\bottomrule
\end{tabular}

}

\end{minipage}%
\newline
\begin{minipage}[t]{\linewidth}
\subcaption{\label{tbl-vae}Variational Autoencoder}

{\centering

\begin{tabular}[t]{lll}
\toprule
 & Hidden Dim. & Epochs\\
\midrule
Synthetic & 2 & 100\\
Real-World & 8 & 250\\
\bottomrule
\end{tabular}

}

\end{minipage}%

\end{table}

\hypertarget{sec-empirical-2}{%
\section{Experiments}\label{sec-empirical-2}}

\hypertarget{endogenous-macrodynamics}{%
\subsection{Endogenous Macrodynamics}\label{endogenous-macrodynamics}}

\hypertarget{potential-mitigation-strategies}{%
\subsection{Potential Mitigation
Strategies}\label{potential-mitigation-strategies}}

\hypertarget{sec-empirical-2-mitigate}{%
\subsubsection{Gravitational Counterfactual
Explanations}\label{sec-empirical-2-mitigate}}

A straight-forward choice simply extends the baseline approach by
\protect\hyperlink{ref-wachter2017counterfactual}{{[}5{]}}: instead of
only penalizing the distance of the individuals' counterfactual to its
factual, we propose penalizing its distance to some sensible point in
the target domain, for example the sample average: \(\bar{\mathbf{x}}\).
For such a recourse objective, higher choices of \(\lambda_2\) relative
to \(\lambda_1\) will lead counterfactuals to gravitate towards the
specified point in the target domain. In the remainder of this paper we
will therefore refer to this approach as \textbf{Gravitational}
generator, when we investigate its potential usefulness for mitigating
endongenous macrodynamics\footnote{Note that despite the naming
  convention our goal here is not to provide yet another counterfactual
  generator, but merely investigate the most simple penalty we can think
  of with respect to its effectiveness.}.

\hypertarget{a-note-on-convergence}{%
\paragraph{A note on convergence}\label{a-note-on-convergence}}

For this simple mitigating strategy underlying the Gravitational
generator to work as expected, one needs to ensure that counterfactual
search continues, even after a predetermined threshold probability
\(\gamma\) has potentially already been reached.
Figure~\ref{fig-convergence} illustrates this distinction: if one
chooses to terminate search once the desired threshold is reached (left
panel) the gravitational pull towards \(\bar{\mathbf{x}}\) is never
actually satisfied (compare to right panel). More generally, if
convergence is defined simply in terms of flipping the predicted label
with some desired degree of confidence, this corresponds to essentially
ignoring any parts of the counterfactual search objective that do not
involve \(\ell(M(f(s_k^\prime)),t)\) beyond that point. While this may
be appropriate for some applications, in general this seems like an odd
convention. Since we nonetheless seen convergence specified simply in
terms of reaching the threshold probability in some places\footnote{\protect\hyperlink{ref-joshi2019towards}{{[}8{]}}
  define convergence of Algorithm 1 in this way. The implementation of
  \protect\hyperlink{ref-wachter2017counterfactual}{{[}5{]}} in CARLA is
  also defined in this way.}, we thought it worth making this
distinction explicit.

\begin{figure}

{\centering \includegraphics[width=0.45\textwidth,height=\textheight]{www/gravitational_generator_comparison.png}

}

\caption{\label{fig-convergence}Comparison of counterfactual search
outcome with simple (left) and strict convergence (right).}

\end{figure}

\hypertarget{sec-discussion}{%
\section{Discussion}\label{sec-discussion}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Shift of focus from individual to group of individuals (related:
  https://www.researchgate.net/publication/353073138\_Generating\_Collective\_Counterfactual\_Explanations\_in\_Score-Based\_Classification\_via\_Mathematical\_Optimization)
\item
  Convergence criterium matters: terminating once threshold probability
  is reached may not be optimal (see e.g.~REVISE)
\item
  Optimizer choice matters: dimensionality is typically low, so no
  obvious benefit to using ADAM.

  \begin{itemize}
  \tightlist
  \item
    This might be better placed in JuliaCon proceedings, perhaps backed
    by small blog post on the matter.
  \end{itemize}
\item
  Mitigating strategy: penaliye distance from centroid.
\end{enumerate}

\hypertarget{sec-limit}{%
\section{Limitations and Future Work}\label{sec-limit}}

While we believe that this work constitutes a valuable starting point
for addressing existing issues in Algorithmic Recourse from a fresh
perspective, we are aware of several of its limitations. In the
following we highlight some of these limitations and point to avenues
for future research.

\hypertarget{experimental-setup}{%
\subsection{Experimental Setup}\label{experimental-setup}}

The experimental setup proposed here is designed to mimic a real-world
recourse process in a simple fashion. In practice, models are in fact
updated on a regular basis
\protect\hyperlink{ref-upadhyay2021towards}{{[}15{]}}. We also find it
plausible to assume that the implementation of recourse happens
periodically for different individuals, rather that all at once at time
\(t=0\). That being said, our experimental design is a vast
over-simplification of potential real-world scenarios. In practice, any
endogenous shifts that may occur can be expected to be entangled with
exogenous shifts of the nature investigated in
\protect\hyperlink{ref-upadhyay2021towards}{{[}15{]}}. We also make
implicit assumptions about the utility functions of the involved agents
that may well be too simple: individuals seeking recourse are assumed to
always implement the proposed Counterfactual Explanations; conversely,
the agent in charge of the model \(M\) is assumed to always treat
individuals that have implemented valid recourse as if they were truly
now in the target class. Relating this back to the consumer credit
example, we assume that the would-be borrowers are always willing and
able to implement recourse and the bank is always willing to provide
credit as would-be borrowers move across the decision boundary. In
practice it is doubtful that agents behave according to such simple
rules. Nonetheless, we think that our simple framework offers a starting
point for future work on recourse dynamics (both endogenous and
exogenous dynamics).

\hypertarget{sec-limit-data}{%
\subsection{Data}\label{sec-limit-data}}

Largely in line with the existing literature on Algorithmic Recourse, we
have limited our analysis of real-world data to three commonly used
benchmark datasets that involve binary prediction tasks. Future work may
benefit from including novel datasets or extending the analysis to
multi-class or regression problems, the latter arguably representing the
most common objective in Finance and Economics. It is also worth
mentioning that the use of real-world datasets considered in this work
is constrained by the fact that at the time of writing
\texttt{CounterfactualExplanations.jl} only supports continuous
features, at least of some of the counterfactual generators considered
here. The fact that we therefore had to discard discrete features led to
relatively poor initial performance of our classifiers in some cases.
While this is indeed a limitation we intend to address in future and
derivative work, our findings with respect to endogenous macrodynamics
do not hinge on strong classifier performance.

\hypertarget{classifiers}{%
\subsection{Classifiers}\label{classifiers}}

For reasons stated earlier we have limited our analysis to
differentiable linear and non-linear classifiers, in particular logistic
regression and deep neural networks. While these sorts of classifiers
have also typically been analyzed in the existing literature on
Counterfactual Explanations and Algorithmic Recourse, they represent
only a subset of popular machine learning models employed in practice -
both black-box and glass-box. Despite the success and popularity of deep
learning in the context of high-dimensional data such as image, audio
and video, empirical evidence suggests that other models such as boosted
decision trees may have an edge when it comes to lower-dimensional
tabular datasets, such as the ones considered here
\protect\hyperlink{ref-grinsztajn2022tree}{{[}36{]}}.

\hypertarget{sec-conclusion}{%
\section{Concluding Remarks}\label{sec-conclusion}}

\hypertarget{acknowledgment}{%
\section*{Acknowledgment}\label{acknowledgment}}
\addcontentsline{toc}{section}{Acknowledgment}

P. A. thanks \ldots{}

\pagebreak
\FloatBarrier

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\leavevmode\vadjust pre{\hypertarget{ref-borch2022machine}{}}%
\CSLLeftMargin{{[}1{]} }%
\CSLRightInline{C. Borch, {``Machine learning, knowledge risk, and
principal-agent problems in automated trading,''} \emph{Technology in
Society}, p. 101852, 2022.}

\leavevmode\vadjust pre{\hypertarget{ref-o2016weapons}{}}%
\CSLLeftMargin{{[}2{]} }%
\CSLRightInline{C. O'neil, \emph{Weapons of math destruction: How big
data increases inequality and threatens democracy}. Crown, 2016.}

\leavevmode\vadjust pre{\hypertarget{ref-rudin2019stop}{}}%
\CSLLeftMargin{{[}3{]} }%
\CSLRightInline{C. Rudin, {``Stop explaining black box machine learning
models for high stakes decisions and use interpretable models
instead,''} \emph{Nature Machine Intelligence}, vol. 1, no. 5, pp.
206--215, 2019.}

\leavevmode\vadjust pre{\hypertarget{ref-pawelczyk2021carla}{}}%
\CSLLeftMargin{{[}4{]} }%
\CSLRightInline{M. Pawelczyk, S. Bielawski, J. van den Heuvel, T.
Richter, and G. Kasneci, {``Carla: A python library to benchmark
algorithmic recourse and counterfactual explanation algorithms,''}
\emph{arXiv preprint arXiv:2108.00783}, 2021.}

\leavevmode\vadjust pre{\hypertarget{ref-wachter2017counterfactual}{}}%
\CSLLeftMargin{{[}5{]} }%
\CSLRightInline{S. Wachter, B. Mittelstadt, and C. Russell,
{``Counterfactual explanations without opening the black box: Automated
decisions and the GDPR,''} \emph{Harv. JL \& Tech.}, vol. 31, p. 841,
2017.}

\leavevmode\vadjust pre{\hypertarget{ref-altmeyer2022CounterfactualExplanations}{}}%
\CSLLeftMargin{{[}6{]} }%
\CSLRightInline{P. Altmeyer, \emph{{CounterfactualExplanations.jl - a
Julia package for Counterfactual Explanations and Algorithmic
Recourse}}. 2022. Available:
\url{https://github.com/pat-alt/CounterfactualExplanations.jl}}

\leavevmode\vadjust pre{\hypertarget{ref-schut2021generating}{}}%
\CSLLeftMargin{{[}7{]} }%
\CSLRightInline{L. Schut \emph{et al.}, {``Generating interpretable
counterfactual explanations by implicit minimisation of epistemic and
aleatoric uncertainties,''} in \emph{International conference on
artificial intelligence and statistics}, 2021, pp. 1756--1764.}

\leavevmode\vadjust pre{\hypertarget{ref-joshi2019towards}{}}%
\CSLLeftMargin{{[}8{]} }%
\CSLRightInline{S. Joshi, O. Koyejo, W. Vijitbenjaronk, B. Kim, and J.
Ghosh, {``Towards realistic individual recourse and actionable
explanations in black-box decision making systems,''} \emph{arXiv
preprint arXiv:1907.09615}, 2019.}

\leavevmode\vadjust pre{\hypertarget{ref-mothilal2020explaining}{}}%
\CSLLeftMargin{{[}9{]} }%
\CSLRightInline{R. K. Mothilal, A. Sharma, and C. Tan, {``Explaining
machine learning classifiers through diverse counterfactual
explanations,''} in \emph{Proceedings of the 2020 conference on
fairness, accountability, and transparency}, 2020, pp. 607--617.}

\leavevmode\vadjust pre{\hypertarget{ref-antoran2020getting}{}}%
\CSLLeftMargin{{[}10{]} }%
\CSLRightInline{J. Antorán, U. Bhatt, T. Adel, A. Weller, and J. M.
Hernández-Lobato, {``Getting a clue: A method for explaining uncertainty
estimates,''} \emph{arXiv preprint arXiv:2006.06848}, 2020.}

\leavevmode\vadjust pre{\hypertarget{ref-karimi2020survey}{}}%
\CSLLeftMargin{{[}11{]} }%
\CSLRightInline{A.-H. Karimi, G. Barthe, B. Schölkopf, and I. Valera,
{``A survey of algorithmic recourse: Definitions, formulations,
solutions, and prospects,''} \emph{arXiv preprint arXiv:2010.04050},
2020.}

\leavevmode\vadjust pre{\hypertarget{ref-verma2020counterfactual}{}}%
\CSLLeftMargin{{[}12{]} }%
\CSLRightInline{S. Verma, J. Dickerson, and K. Hines, {``Counterfactual
explanations for machine learning: A review,''} \emph{arXiv preprint
arXiv:2010.10596}, 2020.}

\leavevmode\vadjust pre{\hypertarget{ref-ustun2019actionable}{}}%
\CSLLeftMargin{{[}13{]} }%
\CSLRightInline{B. Ustun, A. Spangher, and Y. Liu, {``Actionable
recourse in linear classification,''} in \emph{Proceedings of the
conference on fairness, accountability, and transparency}, 2019, pp.
10--19.}

\leavevmode\vadjust pre{\hypertarget{ref-karimi2021algorithmic}{}}%
\CSLLeftMargin{{[}14{]} }%
\CSLRightInline{A.-H. Karimi, B. Schölkopf, and I. Valera,
{``Algorithmic recourse: From counterfactual explanations to
interventions,''} in \emph{Proceedings of the 2021 ACM conference on
fairness, accountability, and transparency}, 2021, pp. 353--362.}

\leavevmode\vadjust pre{\hypertarget{ref-upadhyay2021towards}{}}%
\CSLLeftMargin{{[}15{]} }%
\CSLRightInline{S. Upadhyay, S. Joshi, and H. Lakkaraju, {``Towards
robust and reliable algorithmic recourse,''} \emph{arXiv preprint
arXiv:2102.13620}, 2021.}

\leavevmode\vadjust pre{\hypertarget{ref-carrizosa2021generating}{}}%
\CSLLeftMargin{{[}16{]} }%
\CSLRightInline{E. Carrizosa, J. Ramırez-Ayerbe, and D. Romero,
{``Generating collective counterfactual explanations in score-based
classification via mathematical optimization,''} 2021.}

\leavevmode\vadjust pre{\hypertarget{ref-rabanser2019failing}{}}%
\CSLLeftMargin{{[}17{]} }%
\CSLRightInline{S. Rabanser, S. Günnemann, and Z. Lipton, {``Failing
loudly: An empirical study of methods for detecting dataset shift,''}
\emph{Advances in Neural Information Processing Systems}, vol. 32,
2019.}

\leavevmode\vadjust pre{\hypertarget{ref-chandola2009anomaly}{}}%
\CSLLeftMargin{{[}18{]} }%
\CSLRightInline{V. Chandola, A. Banerjee, and V. Kumar, {``Anomaly
detection: A survey,''} \emph{ACM computing surveys (CSUR)}, vol. 41,
no. 3, pp. 1--58, 2009.}

\leavevmode\vadjust pre{\hypertarget{ref-widmer1996learning}{}}%
\CSLLeftMargin{{[}19{]} }%
\CSLRightInline{G. Widmer and M. Kubat, {``Learning in the presence of
concept drift and hidden contexts,''} \emph{Machine learning}, vol. 23,
no. 1, pp. 69--101, 1996.}

\leavevmode\vadjust pre{\hypertarget{ref-gama2014survey}{}}%
\CSLLeftMargin{{[}20{]} }%
\CSLRightInline{J. Gama, I. Žliobaitė, A. Bifet, M. Pechenizkiy, and A.
Bouchachia, {``A survey on concept drift adaptation,''} \emph{ACM
computing surveys (CSUR)}, vol. 46, no. 4, pp. 1--37, 2014.}

\leavevmode\vadjust pre{\hypertarget{ref-nelson2015evaluating}{}}%
\CSLLeftMargin{{[}21{]} }%
\CSLRightInline{K. Nelson, G. Corbin, M. Anania, M. Kovacs, J. Tobias,
and M. Blowers, {``Evaluating model drift in machine learning
algorithms,''} in \emph{2015 IEEE symposium on computational
intelligence for security and defense applications (CISDA)}, 2015, pp.
1--8.}

\leavevmode\vadjust pre{\hypertarget{ref-ackerman2021machine}{}}%
\CSLLeftMargin{{[}22{]} }%
\CSLRightInline{S. Ackerman, P. Dube, E. Farchi, O. Raz, and M.
Zalmanovici, {``Machine learning model drift detection via weak data
slices,''} in \emph{2021 IEEE/ACM third international workshop on deep
learning for testing and testing for deep learning (DeepTest)}, 2021,
pp. 1--8.}

\leavevmode\vadjust pre{\hypertarget{ref-de2021framework}{}}%
\CSLLeftMargin{{[}23{]} }%
\CSLRightInline{R. M. B. de Oliveira and D. Martens, {``A framework and
benchmarking study for counterfactual generating methods on tabular
data,''} \emph{Applied Sciences}, vol. 11, no. 16, p. 7274, 2021.}

\leavevmode\vadjust pre{\hypertarget{ref-dombrowski2021diffeomorphic}{}}%
\CSLLeftMargin{{[}24{]} }%
\CSLRightInline{A.-K. Dombrowski, J. E. Gerken, and P. Kessel,
{``Diffeomorphic explanations with normalizing flows,''} 2021.}

\leavevmode\vadjust pre{\hypertarget{ref-pindyck2014microeconomics}{}}%
\CSLLeftMargin{{[}25{]} }%
\CSLRightInline{R. S. Pindyck and D. L. Rubinfeld,
\emph{Microeconomics}. Pearson Education, 2014.}

\leavevmode\vadjust pre{\hypertarget{ref-arcones1992bootstrap}{}}%
\CSLLeftMargin{{[}26{]} }%
\CSLRightInline{M. A. Arcones and E. Gine, {``On the bootstrap of u and
v statistics,''} \emph{The Annals of Statistics}, pp. 655--674, 1992.}

\leavevmode\vadjust pre{\hypertarget{ref-hanneke2007bound}{}}%
\CSLLeftMargin{{[}27{]} }%
\CSLRightInline{S. Hanneke, {``A bound on the label complexity of
agnostic active learning,''} in \emph{Proceedings of the 24th
international conference on machine learning}, 2007, pp. 353--360.}

\leavevmode\vadjust pre{\hypertarget{ref-gmsc_data}{}}%
\CSLLeftMargin{{[}28{]} }%
\CSLRightInline{K. Competition, {``Give me some credit, improve on the
state of the art in credit scoring by predicting the probability that
somebody will experience financial distress in the next two years.''}
\url{https://www.kaggle.com/c/GiveMeSomeCredit}}

\leavevmode\vadjust pre{\hypertarget{ref-yeh2009comparisons}{}}%
\CSLLeftMargin{{[}29{]} }%
\CSLRightInline{I.-C. Yeh and C. Lien, {``The comparisons of data mining
techniques for the predictive accuracy of probability of default of
credit card clients,''} \emph{Expert systems with applications}, vol.
36, no. 2, pp. 2473--2480, 2009.}

\leavevmode\vadjust pre{\hypertarget{ref-pedregosa2011scikit}{}}%
\CSLLeftMargin{{[}30{]} }%
\CSLRightInline{F. Pedregosa \emph{et al.}, {``Scikit-learn: Machine
learning in python,''} \emph{the Journal of machine Learning research},
vol. 12, pp. 2825--2830, 2011.}

\leavevmode\vadjust pre{\hypertarget{ref-pace1997sparse}{}}%
\CSLLeftMargin{{[}31{]} }%
\CSLRightInline{R. K. Pace and R. Barry, {``Sparse spatial
autoregressions,''} \emph{Statistics \& Probability Letters}, vol. 33,
no. 3, pp. 291--297, 1997.}

\leavevmode\vadjust pre{\hypertarget{ref-carlisle2019racist}{}}%
\CSLLeftMargin{{[}32{]} }%
\CSLRightInline{C. M., {``Racist data destruction? - a boston housing
dataset controversy,''} 2019, Available:
\url{https://medium.com/@docintangible/racist-data-destruction-113e3eff54a8}}

\leavevmode\vadjust pre{\hypertarget{ref-innes2018fashionable}{}}%
\CSLLeftMargin{{[}33{]} }%
\CSLRightInline{M. Innes \emph{et al.}, {``Fashionable modelling with
flux,''} \emph{arXiv preprint arXiv:1811.01457}, 2018.}

\leavevmode\vadjust pre{\hypertarget{ref-lakshminarayanan2016simple}{}}%
\CSLLeftMargin{{[}34{]} }%
\CSLRightInline{B. Lakshminarayanan, A. Pritzel, and C. Blundell,
{``Simple and scalable predictive uncertainty estimation using deep
ensembles,''} \emph{arXiv preprint arXiv:1612.01474}, 2016.}

\leavevmode\vadjust pre{\hypertarget{ref-borisov2021deep}{}}%
\CSLLeftMargin{{[}35{]} }%
\CSLRightInline{V. Borisov, T. Leemann, K. Seßler, J. Haug, M.
Pawelczyk, and G. Kasneci, {``Deep neural networks and tabular data: A
survey,''} \emph{arXiv preprint arXiv:2110.01889}, 2021.}

\leavevmode\vadjust pre{\hypertarget{ref-grinsztajn2022tree}{}}%
\CSLLeftMargin{{[}36{]} }%
\CSLRightInline{L. Grinsztajn, E. Oyallon, and G. Varoquaux, {``Why do
tree-based models still outperform deep learning on tabular data?''}
\emph{arXiv preprint arXiv:2207.08815}, 2022.}

\end{CSLReferences}

\pagebreak

\hypertarget{tables}{%
\section{Tables}\label{tables}}

\ldots{}

\pagebreak

\hypertarget{figures}{%
\section{Figures}\label{figures}}

\ldots{}

\pagebreak

\hypertarget{code}{%
\section{Code}\label{code}}

\ldots{}

\end{document}