### 2022-08-15 dev/paper/sections/conclusion.qmd

# Concluding Remarks {#sec-conclusion}

This work has revisited and extended some of the most general and defining concepts underlying the literature on Counterfactual Explanations and, in particular, Algorithmic Recourse. We demonstrate that long-held beliefs as to what defines optimality in AR, are too short-sighted to serve as a foundation for applications of recourse in practice. Specifically, we run multiple experiments that simulate the application of recourse in practice using various popular counterfactual generators and find that all of them induce substantial domain and model shifts. We argue that these shifts should be considered as an expected external cost of individual recourse and call for a paradigm shift from individual to collective recourse. By proposing an adapted counterfactual search objective that incorporates this cost, we make that paradigm shift explicit. We show that this modified objective lends itself to mitigation strategies that can be used to effectively decrease the magnitude of induced domain and model shifts. Through our work we hope to inspire future research on this important topic. To this end we have open-sourced all of our code along with a Julia package - `AlgorithmicRecourseDynamics.jl`. The package is built on top of `CounterfactualExplanations.jl` and inherits its extensibility [@altmeyer2022CounterfactualExplanations]. That is to say that future researchers should find it relatively easy to replicate, modify and extend the simulation experiments presented here and apply to their own custom counterfactual generators.