### 2022-08-17 paper/appendix.qmd
uh

---
title: Supplementary Appendix
format:
    html: 
        self-contained: true
        code-fold: true
        execute:
            echo: true
            eval: false
            warning: false
jupyter: julia-1.7
---

This is a supplementary appendix to the research paper **Endogenous Macrodynamics in Algorithmic Recourse**. It contains all of the experimental results, including those not highlighted in the actual paper. It also contains additional information about the proposed counterfactual generators.

# Experimental Results {#sec-results}

{{< include ../dev/notebooks/experiments/synthetic.qmd >}}

{{< include ../dev/notebooks/experiments/real_world.qmd >}}

{{< include ../dev/notebooks/experiments/mitigation_strategies.qmd >}}

# Generators {#sec-generators}

{{< include ../dev/notebooks/generators/gravitational_generator.qmd >}}

{{< include ../dev/notebooks/generators/clap_roar_generator.qmd >}}