### 2022-12-16 docs/src/paper/index.qmd
moving into docs

---
title: Endogenous Macrodynamics in Algorithmic Recourse
subtitle: Research Paper
---

```@meta
CurrentModule = ConformalPrediction
```

This part of the documentation is specific to the research paper "Endogenous Macrodynamics in Algorithmic Recourse". To allow for reproducibility of the results at the time of publication, there is a specific package environment linked to this part of the documentation that **will not be updated**. It can be activated as follows:

```{julia}
using Pkg; Pkg.activate("docs/src/paper")
```

The status of this package environment is shown below:

```{julia}
#| output: true

Pkg.status()
```