### 2022-12-19 docs/src/paper/appendix.qmd
verifying that results check out with updated package

---
title: Supplementary Appendix
---

This is a supplementary appendix to the research paper **Endogenous Macrodynamics in Algorithmic Recourse**. It contains all of the experimental results, including those not highlighted in the actual paper. It also links to additional information about the proposed mitigation strategies.

## Experimental Results {#sec-results}

{{< include experiments/_synthetic.qmd >}}

{{< include experiments/_real_world.qmd >}}

{{< include experiments/_mitigation_strategies.qmd >}}

## Generators {#sec-generators}

The generators have been moved into [`CounterfactualExplanations.jl`](https://github.com/pat-alt/CounterfactualExplanations.jl). For more information so the package [documentation](https://pat-alt.github.io/CounterfactualExplanations.jl/dev).