### 2022-12-23 README.qmd
cleaned up a bit

---
format: 
  commonmark:
    variant: -raw_html
    wrap: none
execute: 
  freeze: auto
  echo: true
  eval: true
  output: false
crossref:
  fig-prefix: Figure
  tbl-prefix: Table
bibliography: bib.bib
jupyter: julia-1.8
---

[![Stable](https://img.shields.io/badge/docs-stable-blue.svg)](https://pat-alt.github.io/AlgorithmicRecourseDynamics.jl/stable) 
[![Dev](https://img.shields.io/badge/docs-dev-blue.svg)](https://pat-alt.github.io/AlgorithmicRecourseDynamics.jl/dev) 
[![Build Status](https://github.com/pat-alt/AlgorithmicRecourseDynamics.jl/actions/workflows/CI.yml/badge.svg?branch=main)](https://github.com/pat-alt/AlgorithmicRecourseDynamics.jl/actions/workflows/CI.yml?query=branch%3Amain) 
[![Coverage](https://codecov.io/gh/pat-alt/AlgorithmicRecourseDynamics.jl/branch/main/graph/badge.svg)](https://codecov.io/gh/pat-alt/AlgorithmicRecourseDynamics.jl) 
[![Code Style: Blue](https://img.shields.io/badge/code%20style-blue-4495d1.svg)](https://github.com/invenia/BlueStyle) 
[![ColPrac: Contributor’s Guide on Collaborative Practices for Community Packages](https://img.shields.io/badge/ColPrac-Contributor's%20Guide-blueviolet.png)](https://github.com/SciML/ColPrac) 
[![Twitter Badge](https://img.shields.io/twitter/url/https/twitter.com/paltmey.svg?style=social&label=Follow%20%40paltmey)](https://twitter.com/paltmey)

# AlgorithmicRecourseDynamics

{{< include docs/src/_intro.qmd >}}