### 2022-12-23 docs/src/_intro.qmd

```{julia}
#| echo: false
include("docs/setup_docs.jl")
eval(setup_docs)
```

`AlgorithmicRecourseDynamics.jl` is a small package for modeling Algorithmic Recourse Dynamics. It builds on `CounterfactualExplanations`, a package for generating counterfactual explanations.

## Basic Usage

Below we first generate some synthetic data for a binary classification task and instantiate an instance of `CounterfactualData`.

```{julia}
N = 1000
xmax = 2
X, ys = make_blobs(
    N, 2; 
    centers=2, as_table=false, center_box=(-xmax => xmax), cluster_std=0.1
)
ys .= ys.==2
X = X'
counterfactual_data = CounterfactualData(X,ys')
```

We then define a simple model for the task and prepare it for use with `CounterfactualExplanations.jl`:

```{julia}
n_epochs = 100
model = Chain(Dense(2,1))
mod = FluxModel(model)
```

To generate algorithmic recourse, we will use a simple generic generator:

```{julia}
generator = GenericGenerator()
```

Finally we train our model on a subset of the data. The chart below shows the results.

```{julia}
#| output: true

data_train, data_test = Data.train_test_split(counterfactual_data)
Models.train(mod, data_train; n_epochs=n_epochs)
plt_original = plot(mod, counterfactual_data; zoom=0, colorbar=false)
display(plt_original)
```

### Simulation

To model the dynamics of algorithmic recourse, we use simulations, in which we repeatedly select as subset of individuals from the non-target class, generate and implement recourse for all of them and finally retrain the model. To set this experiment up, we can use the code below:

```{julia}
models = Dict(:mymodel => mod)
generators = Dict(:wachter => generator)
experiment = set_up_experiment(data_train, data_test, models, generators)
```

Finally, we just run the experiment using default parameter settings that specify the number of rounds, the proportion of individuals to select for recourse and related aspects:

```{julia}
#| output: true

run!(experiment)
```

The chart below shows the data and predictions at the end of the simulation:

```{julia}
#| output: true

new_data = experiment.recourse_systems[1][1].data
new_model = experiment.recourse_systems[1][1].model
plt_original = plot(new_model, new_data; zoom=0, colorbar=false)
```

## Related Research Paper 📝

The package was developed for a research project that investigates the dynamics of various counterfactual generators. You can find the details [here](https://github.com/pat-alt/endogenous-macrodynamics-in-algorithmic-recourse).